### 2014-01-31 creep/creep_yaourt.tex
Balance bibliography Andrade not in intro rheology -> creep xp master curve in caption of fig2 nor on system composition global shear rate and fracture growth related, no causality No discussion of smaller exponents and rework the sentences around SI: GDL4% in Plane-plane

\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4}
\usepackage{graphicx}
\usepackage{dcolumn,xcolor,ulem}
\usepackage{amsmath} 
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage[latin1]{inputenc}
%\usepackage{mathrsfs}

\newcommand{\cor}[1]{\textcolor{red}{#1}}
\newcommand{\change}[1]{\textcolor{blue}{#1}}
\newcommand\gp{\dot\gamma}
\newcommand\gl{\gamma_{\rm loc}}
\newcommand\gpm{\dot\gamma_{\rm min}}
\newcommand\taum{\tau_{\rm min}}

\newcommand{\ti}[1]{\textbf{\color{blue}#1}} % remarques de Thibaut
\newcommand{\titi}[1]{\sout{\textbf{\color{blue}#1}}} % remarques de Thibaut

\begin{document}

%\widowpenalty=1000
%\clubpenalty=1000
%\preprint{APS/123-QED}

 \title{Creep and brittle failure of a protein gel under stress}

\author{M.~Leocmach}
\email{mathieu.leocmach@ens-lyon.fr}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{C.~Perge}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{T.~Divoux}
%\email{divoux@crpp-bordeaux.cnrs.fr}
\affiliation{Centre de Recherche Paul Pascal, CNRS UPR 8641 - 115 avenue Schweitzer, 33600 Pessac, France}
\author{S.~Manneville}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}

\date{\today}

\begin{abstract}
Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids and to rupture under an external load. Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is typical of brittle solids: after a power-law creep regime fully accounted for by linear viscoelasticity and homogeneous deformation, fractures nucleate and grow logarithmically perpendicularly to shear up to sudden rupture. A single equation accounting for those two successive processes nicely captures the full rheological response. The failure time follows a decreasing power-law with the applied shear stress strongly reminiscent of the Basquin law of fatigue for solids. These results are in excellent agreement with recent fiber-bundle models that include damage accumulation on elastic fibers and exemplify protein gels as model brittle soft solids.
\end{abstract}

\pacs{}
\maketitle

%Under a constant external load, a solid deforms and may eventually fail. The resulting slow deformation, referred to as {\it creep}, has been investigated over a century on solid materials \cite{Andrade:1910,Nechad:2005,Miguel:2002} but only a handful of papers have dealt with creep on soft solids, i.e. gels made through the reversible association of biopolymers such as proteins or polysaccharides \cite{Higgs:1990,vanVliet:1995}. These materials, whose physical properties play a key role not only in food science, from set-style yogurt or cheese to jelly-based deserts \cite{Mezzenga:2005} but also in biomechanics, biochemistry and microbiology \cite{Viovy:2000,Grinnell:2010}, behave as elastic solids under small deformations but display remarkable nonlinear behavior featuring fractures prior to irreversible rupture \cite{Bonn:1998,Baumberger:2006,Daniels:2007,Brenner:2013}. Irreversibility directly stems from the existence of an external control parameter, namely temperature (resp. pH) in the case of thermoreversible gels (resp. acid-induced gels), which makes the system impossible to shear-rejuvenate after failure, contrary to other soft glassy materials such as emulsions, colloidal gels and glasses \cite{Cloitre:2000,Divoux:2012,Siebenburger:2012}. While great effort has been devoted to designing protein- and polysaccharide-based soft solids with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}, their nonlinear mechanical behavior under a shear stress has only been partially addressed \cite{vanVliet:1995} and several simple issues remain unexplored such as the spatially resolved scenario for creep and fracture or the physical relevance of the analogy with the failure of hard, brittle solids.

Biogels formed through the self-association of polysaccharide coils, collagen, actin filaments or attractive globular proteins play a major role in biochemistry and microbiology \cite{Viovy:2000}, biological networks and cell mechanics \cite{Stricker:2010} as well as in food science \cite{Mezzenga:2005}. These biomaterials all behave as elastic solids under small deformations but display remarkable nonlinear behavior generally featuring stress- or strain-stiffening \cite{Gardel:2004} and fractures prior to irreversible rupture \cite{Bonn:1998,Baumberger:2006}. Irreversibility stems from the existence of an external control parameter, e.g. temperature or pH in the case of thermoreversible or acid-induced gels respectively, which makes such biogels impossible to shear-rejuvenate after failure, contrary to other soft glassy materials such as emulsions, colloidal gels and glasses \cite{Cloitre:2000,Divoux:2012,Siebenburger:2012}. So far huge effort has been devoted to designing protein gels with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}. However, their mechanical behavior deep into the nonlinear regime has only been partially addressed \cite{vanVliet:1995,Pouzot:2006} and several fundamental issues remain unexplored such as the spatially resolved rupture scenario or the physical relevance of the analogy with brittle failure in hard solids.

%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=7cm]{Fig1.eps}
\caption{(color online) Strain response $\gamma(t)$ in a 4\%~wt. casein gel acidified with 1\%~wt. GDL for an imposed shear stress $\sigma=200$ (\textcolor{red!25!black}{$\bullet$}), 300 (\textcolor{red!50!black}{$\blacktriangledown$}), 400 (\textcolor{red!75!black}{$\blacksquare$}), 550 (\textcolor{red}{$\blacktriangle$}), and 1000~Pa (\textcolor{orange!50!red}{$\blacklozenge$}) from right to left. The gap width is 1~mm. Gray dashes show $\gamma=1$. Inset: failure time $\tau_f$ vs $\sigma$. The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ and $A=40.6\pm 0.5$~s.Pa$^\beta$.
\label{fig1}}
\end{figure} 
%%%%%%%%%%%%

In this Letter we report on stress-induced fracture in protein gels by means of creep experiments coupled to optical and ultrasonic imaging. Gels formed by slow acidification of a sodium caseinate solution \cite{Lucey:1998} display fractures under large strain at fixed low pH values \cite{vanVliet:1995}, which makes them perfect candidates to quantify brittleness in soft solids and tackle the above-mentioned issues. We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: a reversible primary creep regime characterized by a homogeneous strain field and the irreversible nucleation and fracture growth leading to gel failure. Our results are in full agreement with the predictions of some recent fiber-bundle models and hint to universal features of brittle failure common to both soft and hard solids.

%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=7cm,clip]{Fig2.eps}
\caption{(color online) Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize the three successive regimes. $\gpm$ is the minimum shear rate reached at $\taum$ (see text and Suppl. Fig.~2). The yellow line shows the master curve inferred from fitting $\gp(t)$ by Eq.~(\ref{e.fit}) with $\alpha=0.85$, leading to $\lambda=0.378\pm 0.002$ and $\mu=0.187\pm 0.002$. (a)~Primary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in logarithmic scales. Inset: Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 0.1\%. Red lines are power laws $G'\sim G''\sim f^{0.15}$. (b)~Secondary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in linear scales. Gray dashes show the minimum of Eq.~(\ref{e.fit}) reached at $\taum=0.556\tau_f$. Inset: $\taum$ vs $\tau_f$. The red line is $\taum=0.56\tau_f$. (c)~Tertiary creep: $\gp(t)/\gpm$ vs $(\tau_f-t)/\tau_f$ in logarithmic scales with a reversed horizontal axis.\label{fig2}}
\end{figure} 
%%%%%%%%%%%%

Gels are prepared by dissolving sodium caseinate powder (Firmenich) at 4\% wt. in deionized water under gentle mixing at 35$^{\circ}$C and 500~rpm. To induce gelation, 1\% wt. glucono-$\delta$-lactone (GDL) in powder (Firmenich) is dissolved in the solution and its hydrolysis progressively lowers the pH over the course of 8~hours [Fig.~\ref{suppfig1}(a) in the Supplemental Material]. While still liquid, the solution is poured into the gap of a polished Plexiglas Couette cell immersed into a temperature-controlled water tank at 25.0$\pm$0.1$^{\circ}$C \cite{note}. Rheological data are recorded during gel formation by a stress-controlled rheometer through small amplitude oscillatory shear at frequency $f=1$~Hz (Fig.~1 in the Supplemental Material) \cite{note}. Gelation is complete when the elastic ($G'$) and viscous ($G''$) moduli reach a plateau with $G'\gg G''$. A constant stress $\sigma$ is then applied to the sample from time $t=0$ and the subsequent strain response $\gamma(t)$ is monitored. Images of the gel are recorded simultaneously to the rheology (Logitech Webcam Pro 9000). The local velocity and strain fields can also be imaged in the gradient--vorticity plane $(r,z)$ simultaneously to rheology by a custom-made ultrasonic scanner detailed in \cite{Gallot:2013}. In this case, prior to acidification, the sodium caseinate solution is seeded with acoustic tracers here 3\% wt. polyamide spheres (Orgasol 2002 ES3 NAT 3, Arkema, diameter 30~$\mu$m, density 1.02) that do not modify the final gel properties [Fig.~\ref{suppfig1}(b) in the Supplemental Material]. Failure being irreversible, each creep experiment requires to prepare a fresh sample in-situ.

Under a constant applied shear stress $\sigma$, the global strain $\gamma(t)$ displays a robust time dependence (Fig.~\ref{fig1}): $\gamma(t)$ slowly grows with time up to $\gamma \sim 1$ then accelerates until the gel fails at a well-defined time $\tau_f$. These three successive steps are better highlighted in Fig.~\ref{fig2} by focusing on the global shear rate $\gp(t)$. Figures~\ref{fig3} and \ref{fig4} gather the results from {\it local} measurements and are discussed below together with each of the successive regimes inferred from {\it global} data. Note that these failure dynamics do not qualitatively depend on geometry nor on system composition (see Movie~1 in the Supplemental Materials).

%%%%%%%%%%%%
\begin{figure*}[t]
\centering
\includegraphics[width=17.5cm,clip]{Fig3.eps}
\caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes. Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images. Their position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$. The gap width is 2~mm and $r$ is the radial distance to the inner rotating cylinder. Experiment performed on a 4\%~wt. casein gel seeded with 3\%~wt. polyamide spheres and acidified with 1\%~wt. GDL under $\sigma=300$~Pa. See also Movie~2 in the Supplemental Material.
\label{fig3}}
\end{figure*} 
%%%%%%%%%%%%

As seen in the inset of Fig.~\ref{fig1}, the failure time $\tau_f$ sharply decreases as a power-law of $\sigma$ with an exponent $\beta\simeq 5.5$. $\tau_f$ further allows us to rescale all the shear rate data $\gp(t)$ onto the single master curve of Fig.~\ref{fig2}(a) by plotting $\gp/\gp_{\rm min}$ vs $t/\tau_f$, where $\gp_{\rm min}$ is the minimum shear rate reached at a time $\tau_{\rm min}$ [see also Fig.~\ref{suppfig2} in the Supplemental Material for the original $\gp(t)$ data]. On about four decades for $t\lesssim 0.1\tau_f$, the shear rate decreases as a power-law $\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$. This is strongly reminiscent of the {\it primary creep} observed in solids and referred to as Andrade creep \cite{Andrade:1910,Miguel:2002,Nechad:2005}. Interestingly, here, this scaling can be fully accounted for by linear viscoelasticity without invoking any irreversible plasticity. Indeed casein gels display a power-law rheology $G'(f)\sim G''(f)\sim f^{0.15}$ [inset of Fig.~\ref{fig2}(a)], which corresponds to a compliance $J(t)\equiv\gamma(t)/\sigma\sim t^{0.15}$ in the linear deformation regime \cite{Tschoegl:1989}. This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$ and suggests that primary creep is purely reversible without unrecoverable plastic events. Homogeneity and linear response are further confirmed, at least up to the available spatial ($\sim 10~\mu$m) and temporal ($\sim 1$~s) resolutions, by direct visualization which shows no sign of crack or fracture [Fig.~\ref{fig3}(a)] and by ultrasonic imaging which reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)] and no slippage at the Plexiglas walls.

For $0.1\lesssim t/\tau_f\lesssim 0.9$, $\gp(t)$ departs from power-law behavior and goes through a minimum value at time $\tau_{\rm min}= (0.56\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}(b) and inset] as similarly reported for metals \cite{Sundararajan:1989}, solid composite materials \cite{Nechad:2005} and fiber-bundles models (FBMs) \cite{Kovacs:2008,Jagla:2011}. This linearity between $\tau_{\rm min}$ and $\tau_f$, also known as the Monkman-Grant relation \cite{Monkman:1956}, allows one to ``predict'' the failure time from the intermediate-time response. During this {\it secondary creep} regime, regularly-spaced cracks nucleate from the top and bottom edges of the Couette cell and start growing perpendicular to the applied stress [Fig.~\ref{fig3}(b), see also Movie~2 in the Supplemental Material]. These macroscopic fractures are invaded with water expelled from the surrounding gel matrix. Although fractures have not yet entered the ultrasonic region of interest, velocity maps become heterogeneous along the $z$ direction [Fig.~\ref{fig3}(b, left)] and display intermittent fluctuations [Fig~\ref{fig4}(c,d)]. The level of these fluctuations, which can also be seen on the global response for the same applied stress [\textcolor{red!50!black}{$\blacktriangledown$} in Fig.~\ref{fig2}(b)], is poorly reproducible and stress-dependent. We propose that such intermittency arises from microscopic cracks ahead of the fracture tip. Since heterogeneity is much smaller along the radial direction [compare red and black lines in Fig~\ref{fig4}(d)], these ``precursors'' either extend across the full 2-mm gap or induce long-range displacements that span the gap. In any case, such noisy features are typical of slow crack growth in solids \cite{Vanel:2009,Kun:2007} and more generally of avalanche behavior \cite{Sethna:2001}.

In the {\it tertiary creep} regime, for $t\gtrsim 0.9\tau_f$, the shear rate increases by more than four orders of magnitude and diverges as $(\tau_f-t)^{-1}$ [Fig.~\ref{fig2}(c)]. This finite-time singularity corresponds to the final growth of the fractures along the vorticity direction $z$ as they accelerate and eventually meet in the middle of the Couette cell at time $\tau_f$ [Fig.~\ref{fig3}(c-d)]. Ultrasonic velocity maps directly correlate with the crack growth and reveal the complex structure of the displacement field at the tip and around the fracture. In particular, Fig.~\ref{fig3}(d, left) suggests that the fracture is initiated at the inner cylinder at $r=0$ and the presence of large positive and negative velocities in the vicinity of the crack tip is indicative of strong compression and recoil of the gel matrix. Finally, the fracture length $\ell(t)$ is observed to grow logarithmically with $(\tau_f-t)$ upon approaching $\tau_f$ [Fig~\ref{fig4}(d)]. In other words one has ${\rm d}\ell/{\rm d}t\sim\gp(t)\sim (\tau_f-t)^{-1}$, which indicates that the global shear rate is linked to fracture-induced displacements.

%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=17.5cm,clip]{Fig4.eps}
\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ ($\bullet$), $1.7\,10^{-2}$ (\textcolor{blue}{$\blacktriangledown$}), and 0.15 (\textcolor{red}{$\blacksquare$}). Solid lines are linear profiles. (c)~Spatiotemporal diagram of the local velocity $\langle v(r,z,t)\rangle_r$ averaged over the radial direction $r$ and plotted in linear color levels as a function of $z$ and $t/\tau_f$. (d)~Standard deviation $\delta_z v(t)$ of $\langle v(r,z,t)\rangle_r$ taken over the vertical direction $z$ (thick black line) together with corresponding standard deviation $\delta_r v(t)$ computed over the radial direction $r$ on the $z$-average $\langle v(r,z,t)\rangle_z$ (thin red line). (e)~Fracture length $\ell(t)$ vs $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$) and normalized by the height $H$ of the Couette cell. Gray dots show the visualization data for the longest fracture which leads to the failure of the sample at $\tau_f$. Red lines are the best fits $\ell(t)=a+b\log(1-t/\tau_f)$ to the visualization data. Same experiment as in Fig.~\ref{fig3} and Supplemental Movie~2.
\label{fig4}}
\end{figure*} 
%%%%%%%%%%%%

Finally, we emphasize that the (reversible) Andrade-like creep and the (irreversible) crack growth are two  physical processes that effectively superimpose to yield the global rheological response. Indeed, as seen from the yellow line in Fig.~\ref{fig2}, the master curve $\gp(t)/\gp_{\rm min}$ vs $t/\tau_f$ is perfectly fitted by:
%the master curve $\tilde{\gp}(t)=\gp(t)/\gp_{\rm min}$ vs $\tilde{t}=t/\tau_f$ is perfectly fitted by:
\begin{equation}
\frac{\gp(t)}{\gp_{\rm min}}=\lambda \left(\frac{t}{\tau_f}\right)^{-\alpha}+ \frac{\mu}{1-t/\tau_f}\,,
%\tilde{\gp}(\tilde{t})=\lambda \tilde{t}^{-\alpha}+ \frac{\mu}{1-\tilde{t}}\,,
\label{e.fit}
\end{equation}
with only two adjustable parameters $\lambda$ and $\mu$ once $\alpha=0.85$ is fixed. The remarkable collapse of the whole data set to such a simple equation allows us to interpret the secondary creep regime as a mere crossover from creep to crack growth.

Let us now summarize and discuss the most prominent results of this Letter. First we have shown that casein gels display a remarkable failure scenario similar to that of brittle solids and characterized by the same three successive creep regimes. Here, the primary creep and its power-law exponent [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] are fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation. Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Gobeaux:2010,Jaishankar:2013} as well as hard-sphere-like colloidal glasses \cite{Siebenburger:2012}. Protein gels thus appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials. By contrast primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.

The logarithmic fracture growth in the tertiary creep regime constitutes our second important result. Such an evolution is also commonly reported in disordered solid materials displaying brittle rupture and interpreted in the framework of Griffith-like models based on global or local energy barriers \cite{Vanel:2009}. However these approaches all predict exponential scalings for $\tau_f(\sigma)$ while our data are best fitted by the decreasing power-law $\tau_f \sim \sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ [inset of Fig.~\ref{fig1}]. This last key result suggests that thermally activated crack growth is not relevant to the present soft solid. Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Kohout:2000}. Basquin law has also been recently predicted by FBMs that combine elastic fibers with a local yield strain and take into account damage accumulation \cite{Kun:2007,Halasz:2012}. Interestingly assuming damage accumulation to be proportional to $\sigma^\gamma$ directly leads to Basquin law with $\beta=\gamma$ and large values of $\gamma$, typically larger than 5 as found here for $\beta$, lead to macroscopic cracks due to the simultaneous rupture of a large number of fibers \cite{Halasz:2012}. More generally FBMs under compressive or elongational load predict three successive creep regimes exactly alike Fig.~\ref{fig2} with similar proportionality between $\tau_{\rm min}$ and $\tau_f$ and finite-time singularity \cite{Nechad:2005,Jagla:2011}. This highlights the relevance of FBMs in the context of creep in protein gels, whose microstructure indeed appears to be formed of strands \cite{Kalab:1983,Roefs:1990}, but also urges to study FBMs in shear geometries to check whether they would be able to predict fracture growth as observed here.

To conclude, the present time- and space-resolved study shows that protein gels qualify as model, brittle soft solids. Prompted by the remarkable simplicity of Eq.~(\ref{e.fit}) that encompasses Andrade creep, the Monkman-Grant relation and finite-time singularity within a single equation, future modeling will undoubtedly focus on the microscopic ingredients needed to predict quantitatively the Monkman-Grant prefactor and the Basquin exponent $\beta$. The next experimental step consists in a statistical study of the fluctuations associated to crack nucleation and growth as well as a systematic investigation of other systems in order to check for universality in the irreversible creep rupture of soft solids in general and of biogels such as actin, alginate or agar gels in particular. Such a study is expected to have important implications in understanding the behavior of biomaterials under extreme stress conditions.

%which separates into three successive phases: (i)~a primary (or Andrade-like) creep regime during which the strain field is homogeneous, (ii)~a secondary creep during which fractures nucleate and grow logarithmically, and (iii)~a finite-time singularity concomittant with the sudden failure of the gel. The failure time displays a non-trivial power-law scaling with the applied shear-stress that contrasts with the exponential scalings usually reported for solid brittle materials \cite{Vanel:2009}.

% within the spatial resolution of a few microns, the strain field is homogeneous [Fig~\ref{fig4}~(b)]. If plastic events take place in this primary regime, they occur at smaller scale. However, given the agreement between the Andrade exponent and the frequency scaling of the linear elastic modulus discussed above, it is more likely that no or a negligible amount of plastic events take place during this regime, corroborating the optical and ultrasonic measurements.

%Compare to the scalings observed in YSF - in particular microgels display a similar power-law scaling, except that the yield point appears explicitely in the scaling as the critical point in a dynamical phase transition \cite{Divoux:2012}. Propose an explanation or at least raise a point. \cite{Chaudhuri:2013}

%FBM that rely on elastic fibers with a local yield strain (or stress) and introduce a non-linear Eyring rheology or a local coupling lead to a different scaling law than the one we find here for the failure time with the applied stress. In their law appears a ``yield stress" \cite{Kun:2003,Nechad:2005,Kovacs:2008,Jagla:2011} that is not present here.

%different from other soft materials such as transient-network systems \cite{Ligoure:2013}, because truly irreversible. The pH is key to the irreversibility and plays the role of an additional control parameter that, when maintained at values lower than the isoelectric point, prevent the system from healing.

%Comparatively, Yield Stress Fluids (YSF) that display a solid-like behavior at rest, either because of a jammed microstructure or attractive interactions \cite{Sciortino:2005,Bonnecaze:2010}, but do not present such external control parameter, display a similar rheological signature during their stress-induced yielding. Indeed the S-shaped curved pictured in figure~\ref{fig2}~(a) has been also reported in \cite{Bauer:2006,Gibaud:2010,Divoux:2011,Siebenburger:2012,Grenard:2013}. However, their yielding scenario is fundamentally different. Local measurements conducted on both type of soft glassy materials have revealed that no fractures take place in the bulk: the microstructure rather reorganizes and flows in a ductile way. Our study reveals that acid-induced gels behave elastically up to the rupture point and that fractures develop and grow logarithmically, perpendicularly to the shear direction prior to an irreversible failure which makes casein gels model {\it brittle soft solids}.

%and yield stress fluids (YSFs) \cite{Bauer:2006,Divoux:2011,Siebenburger:2012,Grenard:2013} that usually exhibit a power-law creep with smaller exponents (0.3 $\lesssim \alpha \lesssim 0.7$) which are not directly related to linear rheology. Besides, the linear behavior of YSF generally differs from a simple power-law scaling, while in the case of solids

\begin{acknowledgments}
The authors thank T.~Gibaud, S.~Lindstr\"om and N.~Taberlet for fruitful discussions, M.~Kal\'ab for kindly providing us with unobtainable references, and A.~Parker at Firmenich for precious advice and for providing the casein and GDL. This work was funded by the Institut Universitaire de France and by the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement No.~258803. 
\end{acknowledgments}

\begin{thebibliography}{}

\bibitem{Viovy:2000} J.L. Viovy, Rev. Mod. Phys. {\bf 72}, 813--872 (2000); Dorfman et al., Chem. Rev. {\bf 113}, 2584--2667 (2013).

\bibitem{Stricker:2010} J. Stricker, T. Falzone and M.L. Gardel, J. Biomech. {\bf 43},  9--14 (2010); F. Grinnell and W.M. Petroll, Annu. Rev. Cell Dev. Biol. {\bf 26}, 335--361 (2010).

\bibitem{Mezzenga:2005} R. Mezzenga, P. Schurtenberger, A. Burbidge and M. Michel, Nature Mater.  {\bf 4}, 729--740 (2005).

\bibitem{Gardel:2004} M.L. Gardel et al., Science {\bf 304}, 1301--1305 (2004); C. Storm et al., Nature {\bf 435}, 191--194 (2005).

\bibitem{Bonn:1998} D. Bonn et al., Science {\bf 280}, 265--267 (1998).

\bibitem{Baumberger:2006} T. Baumberger, C. Caroli and D. Martina, Nature Mater. {\bf 5}, 552-555 (2006).

%\bibitem{Daniels:2007} K. Daniels et al., Phys. Rev. Lett. {\bf 99}, 124501 (2007).

%\bibitem{Brenner:2013} T. Brenner et al., J. Non-Newtonian fluid Mech. {\bf 196}, 1--7 (2013).

\bibitem{Cloitre:2000} M. Cloitre, R. Borrega and L. Leibler, Phys. Rev. Lett. {\bf 85}, 4819--4822 (2000); V. Viasnoff and F. Lequeux, Phys. Rev. Lett. {\bf 89}, 065701 (2002); Coussot et al., Phys. Rev. Lett. {\bf 88}, 175501 (2002).

\bibitem{Divoux:2012} T. Divoux et al., Soft Matter {\bf 8}, 4151--4164 (2012).

\bibitem{Siebenburger:2012} Siebenb\"urger, M. Ballauff and T. Voigtmann, Phys. Rev.
Lett., {\bf 108}, 255701 (2012).

\bibitem{Dickinson:2006} E. Dickinson, Soft Matter {\bf 2}, 642--652 (2006).

\bibitem{Gibaud:2012} T. Gibaud et al., Faraday Discuss. {\bf 158}, 267--284 (2012).

%\bibitem{1990} P.G. Higgs and S.B. Ross-Murphy, Int. J. Biol. Macromol. {\bf 12}, 233--240 (1990).

\bibitem{vanVliet:1995} T. van Vliet and P. Walstra, Faraday Discuss. {\bf 101}, 359--370 (1995); L.G.B. Bremer et al., Colloids Surf. {\bf 51}, 159--170 (1990).

\bibitem{Pouzot:2006} M. Pouzot, J. Coll. Interf. Sci. {\bf 293} 376--383 (2006).

\bibitem{Lucey:1998} J.A. Lucey and H. Singh, Food Research Int. {\bf 30}, 529--542 (1998).

\bibitem{note} Rheological data shown in Figs.~\ref{fig1} and \ref{fig2} and in Supplemental Figs.~1 and 2 were obtained with an MCR 301 rheometer (Anton Paar) in a Couette cell of height 28~mm and gap 1~mm with an inner rotating cylinder of radius 24~mm. Optical and ultrasonic imaging coupled to rheometry (Figs.~\ref{fig3} and \ref{fig4}) were performed on an AR G2 rheometer (TA Instruments) in a Couette cell of height 60~mm and gap 2~mm with an inner rotating cylinder of radius 23~mm. A detailed, quantitative study of the influence of the gap width on the failure dynamics is left for future work.

\bibitem{Gallot:2013} T. Gallot et al., Rev. Sci. Instrum. {\bf 84}, 045107 (2013).

\bibitem{Andrade:1910} E.N. Da C. Andrade, Proc. R. Soc. London A {\bf 84}, 1--12 (1910).

\bibitem{Miguel:2002} M.C. Miguel, A. Vespignani, M. Zaiser and S. Zapperi, Phys. Rev. Lett. {\bf 89}, 165501 (2002); M.C. Miguel, L. Laurson and M.J. Alava, Eur. Phys. J. B {\bf 64}, 443--450 (2008).

\bibitem{Nechad:2005} H. Nechad et al., Phys. Rev. Lett. {\bf 94}, 045501 (2005); J. Mech. Phys. Solids {\bf 53}, 1099-1127 (2005).

\bibitem{Tschoegl:1989}  N.W. Tschoegl, {\it The Phenomenological Theory of Linear Viscoelastic Behavior} (Springer, Berlin, Germany, 1989).

\bibitem{Sundararajan:1989} G. Sundararajan, Mater. Sci. Eng. A {\bf 112}, 205-214 (1989).

\bibitem{Kovacs:2008} K. Kov\'acs et al., Phys. Rev. E \textbf{77}, 036102 (2008).

\bibitem{Jagla:2011} E.A. Jagla, Phys. Rev. E {\bf 83}, 046119 (2011).

\bibitem{Monkman:1956} F. C. Monkman and N. J. Grant, Proc. Am. Soc. Test.
Mater. {\bf 56}, 593-- (1956).

\bibitem{Vanel:2009} L. Vanel et al., J. Phys. D: Appl. Phys. {\bf 42}, 214007 (2009).

\bibitem{Kun:2007} F. Kun et al., J. Stat. Mech. P02003 (2007); Phys. Rev. Lett. {\bf 100}, 094301 (2008); J. Stat. Mech. P01021 (2009).

\bibitem{Sethna:2001} J.P. Sethna, K.A. Dahmen and C.R. Myers, Nature {\bf 410}, 242--250 (2001).

\bibitem{Gobeaux:2010} F. Gobeaux et al., Soft Matter {\bf 6}, 3769--3777 (2010).

\bibitem{Jaishankar:2013} A. Jaishankar and G.H. McKinley, Proc. R. Soc. A {\bf 469}, 20120284 (2013).

\bibitem{Rosti:2010} J. Rosti, J. Koivisto, L. Laurson, M.J. Alava, Phys. Rev. Lett. {\bf 105}, 100601 (2010).

\bibitem{Basquin:1910} O.H. Basquin, Proceedings of the American Society for Testing and Materials ASTEA, {\bf 10}, 625--630 (1910).

\bibitem{Kohout:2000} J. Kohout, Fatigue Fract. Eng. Mater. Struct. {\bf 23}, 969--977 (2000).

%\bibitem{Suresh:2006} S. Suresh, {\it Fatigue of Materials} (Cambridge University Press, Cambridge, England, 2006).

\bibitem{Halasz:2012} Z. Hal\'asz, Z. Danku and F. Kun, Phys. Rev. E {\bf 85}, 016116 (2012).

\bibitem{Kalab:1983} M. Kal\'ab, P. Allan-Wojtas and B.E. Phipps-Todd, Food Microstruct. {\bf 2}, 51--66 (1983).

\bibitem{Roefs:1990} S.P.F.M. Roefs, A.E.A. De Groot-Mostert and T. Van Vliet, Colloids Surf. {\bf 50}, 141--159 (1990).

%\bibitem{vanVliet:1991} T. van Vliet, H.J.M. van Dijk, P. Zon and P. Walstra, Colloid Polym. Sci. {\bf 269}, 620--627 (1991).

%\bibitem{Bremer:1990} L.G.B. Bremer et al., Colloids Surf. {\bf 51}, 159--170 (1990).

%\bibitem{Grenard:2013} V. Grenard, T. Divoux, N. Taberlet and S. Manneville, Soft Matter (2013) - DOI: 10.1039/C3SM52548A

%\bibitem{Sciortino:2005} F. Sciortino and P. Tartaglia, Advances in Physics {\bf 54}, 471--524 (2005).

%\bibitem{Bonnecaze:2010} R.T. Bonnecaze and M. Cloitre, Adv. Polym. Sci. {\bf 236}, 117--161 (2010).

%\bibitem{Bauer:2006} T. Bauer, J. Oberdisse and L. Ramos, Phys. Rev. Lett. {\bf 97}, 258303 (2006).

%\bibitem{Divoux:2011} T. Divoux, C. Barentin and S. Manneville, Soft Matter {\bf 7}, 8409--8418 (2011).

%\bibitem{Gibaud:2010} T. Gibaud, D. Frelat and S. Manneville, Soft Matter {\bf 6}, 3482--3488 (2010).

%\bibitem{Sprakel:2011} J. Sprakel et al., Phys. Rev. Lett. {\bf 106}, 248303 (2011).

%\bibitem{Kun:2003} F. Kun et al., Phys. Rev. E {\bf 67}, 061802 (2003).

%\bibitem{Ghofraniha:2012} N. Ghofraniha et al., Soft Matter {\bf 8}, 6214 (2012).

%\bibitem{Ligoure:2013} C. Ligoure and S. Mora, Rheol. Acta {\bf 52}, 91-114 (2013).

%\bibitem{Roefs:1986} S.P.F.M. Roefs, {\it Structure of acid casein gels. A study of gels formed after acidification in the cold. PhD Thesis}, Wageningen Agricultural University, The Netherlands (1986).

%\bibitem{Schall:2010} P. Schall and M. van Hecke, Annu. Rev. Fluid Mech. {\bf 42} 67--88 (2010).

%\bibitem{Chaudhuri:2013} P. Chaudhuri and J. Horbach, Phys. Rev. E {\bf 88}, 040301(R) (2013).

\end{thebibliography}

\clearpage
\newpage
\setcounter{figure}{0}

\section*{\large Supplemental material}

\subsection*{Supplemental movies}

Supplemental Movie~1 shows the failure of a 4\%~wt. casein gel acidified with 4\%~wt. GDL in a plate-plate geometry for an imposed shear stress $\sigma=120$~Pa. The plate diameter is 50~mm and the gap width is 1~mm. Fractures grow parallel to the vorticity direction, i.e. along the radial direction in this case. The global rheological response is fully similar to that measured in the Couette geometry.

Supplemental Movie~2 shows the creep experiment analyzed in Figs.~\ref{fig3} and \ref{fig4} and performed under $\sigma=300$~Pa on a 4\%~wt. casein gel seeded with 3\%~wt. polyamide spheres and acidified with 1\%~wt. GDL. The gap width of the Couette cell is 2~mm and its height is 60~mm. Images recorded by a standard webcam (Logitech Webcam Pro 9000) are shown in the top left panel. Velocity maps $v(r,z,t)$ inferred from ultrasonic imaging by averaging over 4~s are shown in the top right panel using linear color levels. The vertical position $z=0$ on the ultrasonic images corresponds to about 20~mm below the top of the Couette cell. The two bottom graphs show the global shear rate response $\gp(t)$ (left) and the strain response $\gamma(t)$ recorded by the rheometer (AR G2, TA Instruments) simultaneously to optical and ultrasonic imaging.

\subsection*{Supplemental figures}

Supplemental Figure~\ref{suppfig1} illustrates the gelation process of a 4\% wt. sodium caseinate solution seeded with 3\% wt. polyamide spheres through time-resolved measurements of the elastic and viscous moduli, respectively $G'$ and $G''$, under small amplitude oscillatory shear. GDL contains an ester function which, once added to the solution at time $t=0$, hydrolyzes spontaneously and leads to a slow decrease of the pH [Fig.~\ref{suppfig1}(a)] towards the casein isoelectric point (${\rm pH}\simeq 4.6$) at which casein particles aggregate. Gelation actually starts earlier at ${\rm pH}\simeq 5$ as shown by gray dashed lines and as already discussed in Ref.~\cite{Lucey:1998supp}. At this point the elastic and viscous moduli display a sudden increase then overshoot and converge towards their steady-state values where $G'\gg G''$ [Fig.~\ref{suppfig1}(b)]. Note that the maximum of the overshoot is reached around the isoelectric point. The decrease of both moduli at lower pH is usually attributed to the over-acidification which enhances the repulsive electrostatic interactions between casein particles of net positive charge \cite{Dickinson:2002supp}.

Supplemental Figure~\ref{suppfig2} shows the shear rate $\gp(t)$ obtained by differentiating the raw strain of Fig.~\ref{fig1} in the main text. For $0.1\lesssim t/\tau_f\lesssim 0.9$, i.e. in the secondary creep regime, $\gp(t)$ is smoothed using a moving average over $\delta t\simeq 0.01\tau_f$ in order to remove high-frequency noise due to differentiation at very small shear rates. The same data normalized by its minimum value $\gp_{\rm min}$ is plotted in Fig.~\ref{fig2}(a) as a function of $t/\tau_f$ leading to a nice collapse onto a single curve whatever the imposed shear stress.

 %%%%%%%%%%%%
\begin{figure}[th]
\centering
\includegraphics[width=7cm,clip]{SuppFig1.eps}
\caption{(color online) Gelation process of a 4\%~wt. casein suspension seeded with 3\%~wt. polyamide spheres and acidified with 1\%~wt. GDL. (a)~pH vs time $t$. (b)~Linear viscoelastic moduli $G'$ (top, in blue) and $G''$ (bottom, in red) recorded as a function of time $t$ for fixed frequency $f=1$~Hz and strain amplitude 0.1\%. Dashed lines indicate that the system starts to build significant elastic properties when the pH decreases below about 5. Lighter colors in (b) correspond to a sample free of polyamide spheres, showing that the addition of acoustic contrast agents does not affect significantly the gelation process and the final viscoelastic properties of the gel.
\label{suppfig1}}
\end{figure}

 %%%%%%%%%%%%
\begin{figure}[h]
\centering
\includegraphics[width=7cm,clip]{SuppFig2.eps}
\caption{(color online) Shear rate response $\gp(t)$ in a 4\%~wt. casein gel acidified with 1\%~wt. GDL for an imposed shear stress $\sigma=200$ (\textcolor{red!25!black}{$\bullet$}), 300 (\textcolor{red!50!black}{$\blacktriangledown$}), 400 (\textcolor{red!75!black}{$\blacksquare$}), 550 (\textcolor{red}{$\blacktriangle$}), and 1000~Pa (\textcolor{orange!50!red}{$\blacklozenge$}) from right to left. The gap width is 1~mm. The yellow line shows the power-law behavior $\gp(t)\sim t^{-0.85}$.
\label{suppfig2}}
\end{figure} 
%%%%%%%%%%%%

\begin{thebibliography}{}

\bibitem{Lucey:1998supp} J.A. Lucey and H. Singh, Food Res. Int. A {\bf 30}, 529--542 (1998).

\bibitem{Dickinson:2002supp} E. Dickinson and L.M. Merino, Food Hydrocolloid {\bf 16}, 321--331 (2002).

\end{thebibliography}

\end{document}