### 2014-05-23 poster_Okinawa.tex
Split G' and pH in groupplot import zooms from plis_figure New take home message

% We switch to portrait mode. This works as advertised.
\documentclass[a0,portrait]{a0poster}
% You might find the 'draft' option to a0 poster useful if you have
% lots of graphics, because they can take some time to process and
% display. (\documentclass[a0,draft]{a0poster})

%\usepackage[utf8]{inputenc}

% Switch off page numbers on a poster, obviously, and section numbers too.
\pagestyle{empty}
\setcounter{secnumdepth}{0}

%fonts
\usepackage[T1]{fontenc}
\usepackage{fontspec}
%\usepackage[oldstylenums, largesmallcaps]{kpfonts}
%\setmainfont[Numbers=OldStyle]{Tex Gyre Pagella}
\setmainfont{Tex Gyre Pagella}
\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}
%\renewcommand*\sfdefault{ugq}

\usepackage{hyperref}
\hypersetup{%
	pdftitle={Wrinkling and nested buckling in a confined protein gel},%the title
	pdfauthor={Mathieu Leocmach},%your name
}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{siunitx}

\usepackage{multirow}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, positioning, intersections, decorations.markings, decorations.shapes, mindmap, shapes.geometric, matrix, patterns}

\usepackage{pgfplots}
%\usepgfplotslibrary{units}
\usepgfplotslibrary{groupplots}
\pgfplotsset{every axis/.append style={xlabel near ticks,ylabel near ticks,mark size={0.2em}}}
\pgfplotsset{every axis plot post/.append style={very thick}}

\usepgfplotslibrary{external}
%\tikzexternalize
%\tikzsetexternalprefix{fig_Rome/}
\tikzset{external/system call={lualatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"}}

\usepackage{ragged2e}
\RaggedRight

\definecolor{Main}{rgb}{1, 0.57, 0}
\definecolor{Accent1}{rgb}{1,0.28,0}
\definecolor{Accent2}{rgb}{1,0.74,0}

% see documentation for a0poster class for the size options here
\let\Textsize\normalsize
\def\Norulehead#1{\noindent\hbox to \hsize{\hfil\LARGE\textcolor{Main}{\textsf{#1}}}\bigskip}
\def\Head#1{\Norulehead{\dotfill #1}}
\def\LHead#1{\noindent{\LARGE #1}\smallskip}
\def\Subhead#1{\noindent{\large\color{Accent1}\textsc{#1}}}
\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{#1}}}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.
\usepackage[absolute,overlay,showboxes
]{textpos}
% Set up the grid
%
% Note that [40mm,40mm] is the margin round the edge of the page --
% it is _not_ the grid size. That is always defined as 
% PAGE_WIDTH/HGRID and PAGE_HEIGHT/VGRID. In this case we use
% 15 x 25. This gives us a wide central column for text (7 grid
% spacings) and two narrow columns (3 each) at each side for 
% pictures, separated by 1 grid spacing.
%
% Note however that texblocks can be positioned fractionally as well,
% so really any convenient grid size can be used.
%
\TPGrid[40mm,40mm]{15}{25}  % 3 - 1 - 7 - 1 - 3 Columns

% Mess with these as you like
\parindent=0pt
%\parindent=1cm
\parskip=0.5\baselineskip

\usepackage{paralist}

%bibliography
\usepackage{natbib}
\usepackage{bibentry}
\def\newblock{\hskip .11em plus .33em minus .07em}

\newlength{\mylength}

%\includeonly{}

\begin{document}
%\tikzset{every mark/.append style={scale=0.8}}
%\pgfplotsset{every axis/.append style={small}}

\bibliographystyle{notitle}
%\nobibliography{sift}

% Understanding textblocks is the key to being able to do a poster in
% LaTeX. In
%
%    \begin{textblock}{wid}(x,y)
%    ...
%    \end{textblock}
%
% the first argument gives the block width in units of the grid
% cells specified above in \TPGrid; the second gives the (x,y)
% position on the grid, with the y axis pointing down.

% You will have to do a lot of previewing to get everything in the 
% right place.

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.
\begin{textblock}{15}(0,0)
\centering
\Title{
Wrinkling of a confined porous layer
}

\smallskip
\LHead{Mathieu Leocmach, Mathieu Nespoulous, Thomas Gibaud, Sébastien Manneville
}\hfill\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}\\

\LHead{\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}\hfill\raisebox{-1.5\baselineskip}{
	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
}}

\end{textblock}

\begin{textblock}{3}(0,2.5)
	\Head{Acid-induced gels}

	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (4\%w \textsc{gdl})

	\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group name=g, group size=1 by 2,
			xticklabels at=edge bottom,
			vertical sep=0,
			},
		xmin=0,xmax=8,
		scale only axis,
		width=\textwidth-4em,
		height=0.3\textwidth,
		extra tick style={grid=major},%
		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.5em}
		]
	\nextgroupplot[
		ymin=0, ymax=7, ylabel={\textcolor{Accent1}{pH}},
		extra y ticks={4.6}, extra y tick labels={},%
		]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
	\node[base left=0] at (axis cs:8,4.6) {isoelectric};

	\nextgroupplot[
		xlabel={time (h)},
		ymin=0, ylabel=\textcolor{Accent2}{$G^\prime$ (\si{\pascal})}
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
	\end{groupplot}
	\end{tikzpicture}

For confocal microscopy we label 10\% of the casein.
\end{textblock}

\begin{textblock}{7}(4,2.5)
	\Head{Sealed cell, spontaneous pattern}

	%\Subhead{Side view}\\
	\tikzsetnextfilename{cell_brushes}
	\tikzset{external/force remake}
	\begin{tikzpicture}
		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\textwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\textwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
		\draw[line width=2pt,Accent1] (0.05\textwidth,-2.5em) -- (0.95\textwidth,-2.5em) (0.05\textwidth,-1pt) -- (0.95\textwidth,-1pt) node[below,pos=0.30, text width=0.4\textwidth] {acrylamide brush ($10\sim 100$ nm)\linebreak $\Rightarrow$ No adhesion};
		\fill[gray] (0,0) rectangle (0.05\textwidth,-2.5em) (\textwidth,0) rectangle (0.95\textwidth,-2.5em) node[pos=1, above left] {spacer};
		\draw[<->] (0.75\textwidth,-2pt) -- (0.75\textwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] (L){$L\sim 2.5\,cm$};
		\node[anchor=north west, inner sep=0] at ($(L.north) - (0.5\textwidth,0)$) {\Subhead{Side view}};
	\end{tikzpicture}

	\Subhead{Top view}\\
	\begin{tikzpicture}[inner sep=0, very thick]
	\setlength{\mylength}{\columnwidth}
	\node[anchor=north west] (a) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2_crop}};
		\node[anchor=north] at (0.44\mylength,0) (b) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6_crop}};
		\node[anchor=north east] at (\mylength,0) (c) {\includegraphics[width=0.4\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
		%zooms
		\node[minimum width = 0.137\mylength, minimum height=0.137\mylength, anchor=north west, draw=Accent2] at ($(a.north west) +(0.111\mylength,-0.072\mylength)$) (bz){};
		\draw[Accent2] (bz.north east) -- (b.north west) (bz.south east) -- (b.south west);
		\node[minimum width = 0.128\mylength, minimum height=0.081\mylength, anchor=north west, draw=Main] at ($(b.north west) +(0.097\mylength,-0.143\mylength)$) (cz){};
		\draw[Main] (cz.north east) -- (c.north west) (cz.south east) -- (c.south west);
		\node[minimum width = 0.156\mylength, minimum height=0.156\mylength, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\mylength,0)$) (dz) {};
		%scale bars
		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.178\mylength,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.176\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};

		%final volumic view
		%\node[anchor=north east] at (\columnwidth,0) (c) {\includegraphics[width=0.4\mylength]{recol_volume_plis}};
	\end{tikzpicture}

\end{textblock}

\begin{textblock}{15}(0,8)
	\Head{Pattern kinetics by transmission macroscope}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}p{2\TPHorizModule}llllll}
	33 min\hfill \begin{tikzpicture}[inner sep=0]
		\draw[ultra thick] (0,0) -- ++(0.355\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{1}{\centi\metre}};
	\end{tikzpicture}&
	 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
	\includegraphics[width=2\TPHorizModule]{prise_0100_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{prise_0130_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{prise_0160_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{prise_0190_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{prise_0250_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{prise_0360_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{prise_0799_resized.jpg}\\
	\end{tabular*}
\end{textblock}

\begin{textblock}{3}(0,10)
	\Subhead{Length scales}
	\begin{itemize}
	\item A fundamental that does not evolve
	\item Nested structures as ``harmonics''
	\end{itemize}
	\Subhead{Qualitatively different from}
	\begin{itemize}
	\item Phase separation patterns 
	\item Turing patterns
	\item Usual wrinkles
	\end{itemize}	 
\end{textblock}

\begin{textblock}{11}(4,11)
	\Head{Confocal Microscopy}
\end{textblock}

\begin{textblock}{3}(4,11)
	\addvspace{2em}
	\Subhead{Procedure}
	\begin{itemize}
		\item Fluorescently label 20\% of the casein
		\item A confocal slice every \SI{5}{\micro\metre}
		\item $(\SI{1.3}{\milli\metre})^2 \times \SI{120}{\micro\metre}$
		\item[$\Rightarrow$] Protein density field
		\item Detect top and bottom surfaces of the gel phase
	\end{itemize}
\end{textblock}

\begin{textblock}{2}(8,11)
	\addvspace{2em}
	\Subhead{Final state}

	\begin{tikzpicture}[inner sep=0]
		\node[above right]{\includegraphics[width=\textwidth]{collage_cut}};
		\draw[ultra thick] ++(0,0.5em) -- ++(0.33\textwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
	\end{tikzpicture}\\
	Tiled, enhanced elevation
\end{textblock}

\begin{textblock}{4}(11,11)
	\addvspace{2em}
	\Subhead{Synæresis and swelling}

	\begin{tikzpicture}
	\begin{axis}[%
		xmin=0, xtick={20,40,...,160},
		ymin=0.5, ymax=1, 
		xlabel={time (min)}, ylabel={$V_\text{gel}/V_\text{cell}$},
		extra tick style={grid=major},%
		extra x ticks={33, 38, 43, 48, 58, 77, 150}, extra x tick labels={},%
		%width=2.65\TPHorizModule,
		width=3.65\TPHorizModule,
		%width=0.8\textwidth, 
		height=1.5\TPVertModule,
		]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60.+3}]{cas4_GDL4_fluo0p4_2.volume};
	\end{axis}
	\end{tikzpicture}
\end{textblock}

\begin{textblock}{15}(0,14)
	\Head{Wrinkling kinetics by confocal microscope}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllll}
	\raisebox{-0.75em}{\begin{tikzpicture}[inner sep=0]
		\node[above right]{\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}};
		\draw[ultra thick] ++(0,0.02\TPHorizModule) -- ++(1.55\TPHorizModule,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
	\end{tikzpicture}}&
	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
\end{tabular*}
\end{textblock}

\begin{textblock}{15}(0,25)
\small{\texttt{The research leading to these results has received funding from the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement n°~258803.}}
\end{textblock}

\textblockcolour{lightgray!50!white}
\TPMargin*{ 0.125\TPHorizModule }
\begin{textblock}{2.875}(12,2.625)%real width of 3
	\Norulehead{Take home}
	\Subhead{Poroelasticity dominates}
	\begin{itemize}
		\item gravity
		\item lubrication
	\end{itemize}
	\Subhead{Two modes}
	\begin{itemize}
		\item blistering
		\item wrinkling
	\end{itemize}
	\Subhead{Patterns}
	\begin{itemize}
		\item self-generated
		\item nested
		\item kinetic wavelength
	\end{itemize}

\end{textblock}%Conclusion

\end{document}