### 2014-10-13 soundbite_palavas.tex
soundbite Palavas

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[francais]{babel}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
\usefonttheme{structuresmallcapsserif}
\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
\usecolortheme{whale}
%\useoutertheme{infolines}

\usecolortheme[named=Accent1]{structure}

\setbeamercolor{alerted text}{fg=Accent2}
%\setbeamercolor{palette primary}{fg=white}
%\setbeamercolor{palette secondary}{bg=Accent1}
%\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}

\setbeamerfont{page number in head/foot}{size=\large}
\setbeamercolor{page number in head/foot}{fg=Main}
% page/total
%\setbeamertemplate{footline}[frame number]
% pas de total
\setbeamertemplate{footline}{%
    	\hfill%
	\usebeamercolor[fg]{page number in head/foot}%
	\usebeamerfont{page number in head/foot}%
	\insertframenumber\kern1em\vskip2pt%
}

\setbeamersize{text margin left=1em}
\setbeamersize{text margin right=1em}

%font
\usepackage[T1]{fontenc}
\usepackage[oldstylenums]{kpfonts}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\pgfplotsset{compat=1.9}
\usepackage{pgfplotstable}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\tikzexternalize
\tikzsetexternalprefix{fig_soundbite_palavas/}
%\tikzset{external/optimize=false}
%\tikzset{external/force remake}

%link or play movies
\usepackage{multimedia}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
\title{Role of the microstructure in the failure of a protein gel under stress}
\author[M. Leocmach]{Mathieu Leocmach}
\date{}
\titlegraphic{
	\begin{tabu}{X[c]X[c]X[c]X[c]}
		\includegraphics[height=3\baselineskip]{Chris}&
		\includegraphics[height=3\baselineskip]{Thomas}&
		\includegraphics[height=3\baselineskip]{Thibaut}&
		\includegraphics[height=3\baselineskip]{Seb}\\
		Christophe Perge & Thomas\linebreak Gibaud & Thibaut\linebreak Divoux & Sebastien Manneville\\
	\end{tabu}

	\vfill
	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
	\includegraphics[height=2\baselineskip]{CRPP}\quad
	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
	}

\newlength{\mylength}

%\includeonly{creep_beamer}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={footnotesize}}

\pgfplotscreateplotcyclelist{earthy}{%
red!40!black,
red!60!black,
red!80!black,
red,
red!80!yellow,
red!60!yellow,
red!40!yellow,
}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{Over-acidified protein gels: beyond isoelectric pH}

	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (\textsc{gdl})\\ %\textit{Bremer et al., Colloids Surf. (1990); van Vliet \& Walstra, Faraday Discuss. (1995); Lucey \& Singh, Food Res. Int. (1997); Leocmach et al. PRL (2014)}

	\smallskip
	\tikzsetnextfilename{prise_cas4}
	\input{fig_soundbite_palavas/prise_cas4.pgf}
	\begin{itemize}
	\item Caseins regain charges, solubility below isoelectric $\Rightarrow$ $G^\prime\searrow$
	\item The sample is still solid
	\end{itemize}
\end{frame}

\begin{frame}{Linear rheology, from power-law to soft glassy}
	\tikzsetnextfilename{sweep_cas4}
	\input{fig_soundbite_palavas/sweep_cas4.pgf}
	\begin{itemize}
	\item Near isoelectric, power law rheology $G^\prime, G^{\prime\prime}$ typical of gels 
	\item Over-acidification lowers exponent and creates a minimum in $G^{\prime\prime}$
	\item Transition to soft-glassy behaviour \\\textit{\footnotesize Sollich, PRE (1998); Mason, Curr. Opin. Colloid Interface Sci. (1999)}
	\end{itemize}
\end{frame}

\begin{frame}{Questions}
\begin{description}
\item[What] is the change in microstructure?
\item[How] does it affects the failure mechanism?
\end{description}
\end{frame}

\begin{frame}{Microscructure: retain network, swell aggregates}
	\tikzsetnextfilename{structure_confocal}
	\input{fig_soundbite_palavas/structure_confocal.pgf}
	\begin{itemize}
	\item Maximum $G^\prime$ corresponds to maximum contrast between phases
	\item Characteristic size $\xi$ increases during over-acidification
	\item Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy
	\end{itemize}
\end{frame}

\begin{frame}{Three creep regimes + softness}
\vfill
	\tikzsetnextfilename{creep_cas4_gdl4}
	\input{fig_soundbite_palavas/creep_cas4_gdl4.pgf}
%	\begin{description}
%		\item[$\tau_f$] divergence of the strain
%		\item[$\tau_\text{min}$] minimum of shear rate $\dot{\gamma}$
%	\end{description}
%	\begin{itemize}
%		\item Erratic position of $\tau_\text{min}$
%		\item Possible to reform bonds? Glassy?
%	\end{itemize}
\begin{itemize}
%\item Failure scenario almost conserved: ``fibre bundle'' network
%\item Other dissipation due to softness: overcompressed gel
\item Prediction of failure from when fractures nucleate more difficult
\end{itemize}

\end{frame}

\begin{frame}[plain]
	\titlepage
\end{frame}

\appendix
\newcounter{finalframe}
\setcounter{finalframe}{\value{framenumber}}

\begin{frame}{Basquin law rescaling}
	\tikzsetnextfilename{basquin}
	\input{fig_soundbite_palavas/basquin.pgf}
\end{frame}

\begin{frame}{Monkman-Grant: Predict failure from nucleation}
	\tikzsetnextfilename{MonkmanGrant}
	\input{fig_soundbite_palavas/MonkmanGrant.pgf}
\end{frame}

\begin{frame}{Wavelength}
	\tikzsetnextfilename{lambda}
	\input{fig_soundbite_palavas/lambda.pgf}
	\textcolor{Accent1}{$\blacktriangleright$} $\lambda\propto e$ compatible with nucleation rate

\end{frame}

\end{document}