### 2014-11-11 plis_figures.tex
Include in tex and generate figures.

\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4-1}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap,fit}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\tikzsetexternalprefix{fig_plis/}
\tikzexternalize
\tikzset{external/force remake}

\definecolor{Main}{rgb}{1, 0.57, 0}
\definecolor{Accent1}{rgb}{1,0.28,0}
\definecolor{Accent2}{rgb}{1,0.74,0}

\begin{document}
\author{Mathieu Leocmach}

\begin{figure}
	\tikzsetnextfilename{dynamics}
	\input{fig_plis/dynamics.pgf}
	\caption{Dynamics of pattern formation. Left: light transmission microscopy. Right: Reconstructed from fluorescent confocal microscopy. Scale bars are \SI{1}{\milli\metre}.}
	\label{fig:dynamics}
\end{figure}

\begin{figure}
	\tikzsetnextfilename{acidification}
	\input{fig_plis/acidification.pgf}
	\caption{Acid-induced protein gels properties behave non-monotonously with pH. (a) pH decrease in 4\%w sodium caseinate solution acidified by 4\%w GDL. Horizontal line indicates isoelectric pH of casen (b) Evolution of storage modulus measured in a rheometer (full adhesion to rotor and stator). (c) Evolution of pore size measured by confocal microscopy in a half coated microscopy cell (no adhesion to cell ceiling, syn\ae{}resis and swelling allowed).}
	\label{fig:acidification}
\end{figure}

\begin{figure}
	\tikzsetnextfilename{sideview}
	\input{fig_plis/sideview.pgf}
	\caption{Modelisation  and measurements. (a) Schematic side view of the cell. The constant thickness gel sheet (orange) is surrounded by water. Yellow lines show the path of transmitted light through the gel phase. A region of interest is highlighted in gray. (b) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling. Scale bar is \SI{100}{\micro\metre} (real size ratio). Evolution of (c) volume of the gel phase relative to cell volume and (d) excess area measured by confocal microscopy.}
	\label{fig:sideview}
\end{figure}

\begin{figure*}
	\tikzsetnextfilename{Darcy_vs_Poiseuille}
	\input{fig_plis/Darcy_vs_Poiseuille.pgf}
	\caption{Comparing model predictions with measured wavelengths. Continuous line is the perfect match ($\lambda_{th}=\lambda_{xp}$), dashed line is the best linear fit through the origin (prefactor is 0.81 in a, 0.75 in b), dotted line is the best affine fit ($\lambda_{xp}=0.56\lambda_{th}+\SI{0.28}{\milli\metre}$).}
	\label{fig:DarcyPoiseuille}
\end{figure*}

\end{document}