### 2014-11-14 plis_figures.tex
permeability : Figure : switch tube 1 and 2 in the sketch, use gray instead of yellow, add a second graph for glycerol 50%, use figure* Supplementary method

\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4-1}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap,fit}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\tikzsetexternalprefix{fig_plis/}
\tikzexternalize
\tikzset{external/force remake}

\definecolor{Main}{rgb}{1, 0.57, 0}
\definecolor{Accent1}{rgb}{1,0.28,0}
\definecolor{Accent2}{rgb}{1,0.74,0}

\begin{document}
\author{Mathieu Leocmach}

\begin{figure}
	\tikzsetnextfilename{dynamics}
	\input{fig_plis/dynamics.pgf}
	\caption{Dynamics of pattern formation. Left: light transmission microscopy. Right: Reconstructed from fluorescent confocal microscopy. Scale bars are \SI{1}{\milli\metre}.}
	\label{fig:dynamics}
\end{figure}

\begin{figure}
	\tikzsetnextfilename{acidification}
	\input{fig_plis/acidification.pgf}
	\caption{Acid-induced protein gels properties behave non-monotonously with pH. (a) pH decrease in 4\%w sodium caseinate solution acidified by 4\%w GDL. Horizontal line indicates isoelectric pH of casen (b) Evolution of storage modulus measured in a rheometer (full adhesion to rotor and stator). (c) Evolution of pore size measured by confocal microscopy in a half coated microscopy cell (no adhesion to cell ceiling, syn\ae{}resis and swelling allowed).}
	\label{fig:acidification}
\end{figure}

\begin{figure}
	\tikzsetnextfilename{sideview}
	\input{fig_plis/sideview.pgf}
	\caption{Modelisation  and measurements. (a) Schematic side view of the cell. The constant thickness gel sheet (orange) is surrounded by water. Yellow lines show the path of transmitted light through the gel phase. A region of interest is highlighted in gray. (b) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling. Scale bar is \SI{100}{\micro\metre} (real size ratio). Evolution of (c) volume of the gel phase relative to cell volume and (d) excess area measured by confocal microscopy.}
	\label{fig:sideview}
\end{figure}

\begin{figure*}
	\tikzsetnextfilename{Darcy_vs_Poiseuille}
	\input{fig_plis/Darcy_vs_Poiseuille.pgf}
	\caption{Comparing model predictions with measured wavelengths. Continuous line is the perfect match ($\lambda_{th}=\lambda_{xp}$), dashed line is the best linear fit through the origin (prefactor is 0.81 in a, 0.75 in b), dotted line is the best affine fit ($\lambda_{xp}=0.56\lambda_{th}+\SI{0.28}{\milli\metre}$).}
	\label{fig:DarcyPoiseuille}
\end{figure*}

\begin{figure}
	\tikzsetnextfilename{mesh}
	\input{fig_plis/mesh.pgf}
	\caption{Gel microstructure. (a) Detail of a confocal micrograph of a 4\%w casein, 4\%w GDL in water, when only the slide is coated (not the cover slip). Time corresponds to the minimum in mesh size. Scale bar is \SI{10}{\micro\metre}. (b) Corresponding radial averaged Fourier transform. Black line is the best fit by Eq.XXX with $\xi=\SI{4.9}{\micro\metre}$ and $d=2.4$. Inset shows the center of the non radial average spectrum (arbitrary logarithmic color scale). Scale bar is \SI{10}{\per\micro\metre}. (c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{1.5}{\micro\metre}$ and $d=2.5$.}
	\label{fig:mesh}
\end{figure}

\begin{figure*}
	\tikzsetnextfilename{permeability}
	\input{fig_plis/permeability.pgf}
	\caption{Permeability measurements. (a) Schematic representation of the experiment. (b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2. Black line is the best exponential fit $Ae^{-t/\tau}$. (b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$. (c) Idem in 50\%w glycerol, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
	\label{fig:permeability}
\end{figure*}

\end{document}