### 2015-02-13 fig_plis/article_plis.tex
chasse au passé Moins de détail sur la perméabilité dans le texte (c'est expliqué en méthode et supplémentary) Explique qu'on est capable d'empècher l'adsorption grâce à l'acrylamide, mais ensuite uniquement "turn off" "turn on" Cite les Video 7 et 8 dans le texte En remet une couche pour l'interprétation de H<>H* Complète le caption de fig:nonmonotonic Protein labelling in Video description

\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,
amsmath,amssymb,prl]{revtex4-1}
\usepackage[utf8]{inputenc}

\usepackage{graphicx} 
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[range-units = single]{siunitx}
%\usepackage{multirow}
\usepackage{bm}
%\usepackage{dcolumn,ulem}
\usepackage{amsfonts}

\definecolor{Main}{rgb}{1, 0.57, 0}
\definecolor{Accent1}{rgb}{1,0.28,0}
\definecolor{Accent2}{rgb}{1,0.74,0}

\newcommand{\seb}[1]{\textbf{\color{blue}#1}} % remarques Seb
\newcommand{\sseb}[1]{\sout{\textbf{\color{blue}#1}}} % remarques Seb

\begin{document}
\title{Hierarchical wrinkling in a confined permeable biogel}
\author{Mathieu Leocmach}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{Mathieu Nespoulous}
\affiliation{Aix-Marseille Université, CNRS, MADIREL UMR 7246, Marseille, France}
\author{Sébastien Manneville}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{Thomas Gibaud}
\email{thomas.gibaud@ens-lyon.fr}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}

\begin{abstract}
The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels -- this is indeed one of the key steps to make yoghurt. Here, using the casein proteins as  building blocks and the acidulent glucono-$\delta$-lactone, we design model thin biogel layers that wrinkles in cascade. Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gel shrinks and swells creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel. We develop an elasto-hydrodynamic model that predicts correctly the wrinkling wavelength. Our results should be universal for acid-induced protein gels as it is based on pH induced charge stabilisation/destabilisation and on control of boundary conditions. As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
\end{abstract}

\maketitle

\section*{Introduction}

%There are many ways and reasons for a film to wrinkle. An elastic film buckles due to excess area with respect to its boundaries. Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}. On the engineering side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}. Such patterns can be triggered by temperature dilation, swelling or the removal of pre-strain. On the biology side wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}. On the physics side analogues of wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}. Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection. Stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002}. Since in practice gravity sets the wavelength in such situation~\cite{Smoluchowski1910}, the wrinkling on a confined viscous substrate has not been studied experimentally.

There are many ways and reasons for a film to wrinkle. An elastic film buckles due to excess area with respect to its boundaries and wrinkles when buckling is hindered by a substrate. The selection of the preferred wrinkling wavelength is peculiar to the situation and is related to the mismatch of elastic properties between the surface and the substrate~\cite{Gough1940, Bijlaard1946, Biot1957, Bowden1998, Cerda2003, Genzer2006}. On the biology side wrinkling-controlled morphogenesis is ubiquitous. Ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}. Difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}. Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}. On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}. Such patterns can be triggered by temperature dilation~\cite{Bowden1998}, swelling~\cite{Hu1998,Kim2010} or the removal of pre-strain~\cite{Genzer2006}.

%ph swelling: hu_1998_Polymer gels with engineered environmentally and kim_2010_Dynamic display of biomolecular patterns
%compressive stress: kim_2011_Hierarchical folding of elastic membranes under

Yet, benchmark experiments that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line, not only from a physics point of view, but also to get fundamental insights into biologically relevant situation. Indeed, stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002} and remains theoretical because, on a free interface, gravity dominates over viscous substrate. In a biological context, thin tissues, e.g. epithelimum or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g. lymph, blood or mucus, a practical situation where a thin viscous substrate would set the wrinkling wavelength. Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.

Here, we introduce model permeable biogels that produce hierarchical wrinkling as a result of their interplay between their gelation dynamics and the confinement conditions. In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins. As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off. Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale. We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates. As shown in Fig.~\ref{fig:dynamics} and in Supplementary Videos 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling. Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates. In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Fig.~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.

\begin{figure}[ht!]
	\includegraphics{dynamics}%
	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w). (a) Light transmission microscopy. Successive generations of patterns are highlighted in colour to stress the absence of growth after formation. (b) 3D reconstruction from fluorescent confocal microscopy. Scale bars are \SI{1}{\milli\metre}.}%
	\label{fig:dynamics}%
\end{figure}

The main goal of the present study is thus to identify the selection mechanism of the wavelength. In the following we first describe the physical chemistry that allows us to form such patterns. Then we  model the situation and identify two possible selection mechanisms, namely porous Darcy flow or viscous Poiseuille flow. Comparing our experimental data to these two models we finally show that a combination of both Darcy and Poiseuille mechanisms accurately describes our data. This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
%Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel film for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel film for short wavelengths.

\section*{System design}

\begin{figure}
	\includegraphics{acidification}%
	\caption{Acid-induced protein gel properties behave non monotonically with pH. (a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL. The horizontal line indicates isoelectric pH of caseins. (b) Corresponding evolution of the fraction of free caseins. Symbols are direct measures upon addition of GDL. Continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert). (c) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator. (d) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e. allowed syn\ae{}resis and swelling (orange dashed line).}%
	\label{fig:acidification}
\end{figure}

To form wrinkled patterns such as the one observed in figure~\ref{fig:dynamics}, ones needs to take advantage of the pH sensitivity of caseins to control the route to gelation, to fine tune the gel properties and the boundary conditions.

The stability of caseins dispersion strongly depend on the pH of the solution. Supplementary Video 3 shows the titration of a caseins dispersion at 1\%w by 1 molar HCl. At pH=7, the caseins are well dispersed then then flocculate between pH=5.5 to 4 and finally at pH=3 the solution become clear again. However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary Video 4. The state diagram of caseins as a function of the pH is related to charge stabilization. At neutral pH caseins are charged negatively and repel each other. As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), caseins molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998}. Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve. In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH (see Supplementary Methods). Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction $x_\text{free}$ is 1. Upon decreasing the pH of the solution to the isolectric point the free casein fraction drop to almost 0, almost all the casein molecule are aggregated. Further decrease of the pH of the solution lead to a partial redissolution of the casein aggregates. At pH=2 the fraction of free casein is 0.35.

%On the Stability of Casein Micelles, Pieter Walstra J. of dairy science 1990

To form a space spanning gel, flocculation needs to be prevented. We use of glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Fig.~\ref{fig:acidification}a. This GDL controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$. As shown on Figure~\ref{fig:acidification}a and b, when adding enough GDL the pH lowers beyond the isoeletric point and $x_\text{free}$ rises again. We monitor the GDL induced gelation with a rheometer in cone plate geometry. As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elastic and viscous modulus of the casein dispersion as a function of time, Figure~\ref{fig:acidification}c. The mechanical properties of the gel are closely related to the fraction of free casein. First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finally the elastic modulus decreases down to 100 Pa. This finite value indicates that the sample is still a gel.

To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy. As shown in Supplementary Figure~\ref{fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}. Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~\ref{fig:permeability}. The fractal model fits the structure at all times but its parameters vary non monotonously during GDL acidification. As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998}. This stiffens the arms of the network and is responsible for the increase of elasticity. Beyond the isoelectric point $\xi$ decreases down to \SI{4}{\micro\metre}. As caseins regain solubility, the strands of the network swell, the pores shrink and the gel weakens.

%XXXCe paragraphe fait-il sens ici ?XXX As seen on the confocal measurements, the gel is porous and is permeable to the solvent.  This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off. To quantify this effect, we make the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}. In the Supplementary Figure~\ref{fig:permeability}, we observe that the aqueous buffer goes through the gel and rises in the tube due to capillary forces. Exploitation of this experiment give the permeability of the gel: \SI{67000}{\square\nano\metre}.

Boundary conditions is the final parameter that is essential to form wrinkled patterns. To turn off the adsorption of the caseins on a surface, we coat it with polymer brushes (see Methods), effectively creating non adhesive boundary conditions for the gel. The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel: $\xi$ increases, the stands of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of the network swell and the gel wants to expend. When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall. In this situation we can monitor $\xi$, Fig.~\ref{fig:acidification}d. $\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.

%We coat the glass slides with acrylamide brushes to prevent the adhesion of the casein. In fig.2c, when the top slide of the optical cell is coated with acrylamide brushes, we then observe that the casein network detaches from the top slide. $\xi$ does not peak at the isoelectric point but remains constant

%Boundary conditions is the final parameter that is essential to form wrinkled patterns. Once the space spanning network is formed and sticks to the container, protein attraction can only proceed by local shrinking of the protein agregates, leading to larger pores.

%The difference in value of $\xi$ between the fully un-coated  and the half-coated optical cell demonstrate that the gel is subject to constrains as the pH decreases. The relaxation of thoses constrain is key to the wrinkling formation when both the top and bottom slide are coated with acrylamide brushes.

%on a change les condition daccroche ce aui permet au contrainte de se relacher

%The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation~\cite{Bremer1989} (Fig.~\ref{fig:acidification}a). Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b). Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplementary Fig.~\ref{fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}. In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{fig:nonmonotonic}).

\section*{Dynamics of the wrinkling process}

Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration favourable to wrinkling (see also Supplementary Videos 7 and 8). Initially the dispersion is homogeneous and fills the entire slit. The black margins correspond to the glass walls. Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation. This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides. We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b). Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods). As expected from the system design, the evolution of the volume is non-monotonous. After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases. Shortly after,  at \SI{36}{\minute}, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area. This confirms that the instability is due to swelling. Around \SI{38}{\minute}, the amplitude of the wrinkles reach the thickness of the slit. The size of the primary pattern is now settled, Supplementary Figures~\ref{fig:patterns} and \ref{fig:growth}. However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21). When the excess volume saturates, the dynamics stops. The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.

Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off. We also confirmed that at any given point in time and space the thickness $h$ of the film of gel is homogeneous. The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is therefore not due to thickness inhomogeneities but to altitude gradients. Indeed the length of the optical path along the vertical direction $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}b.

\begin{figure}
	\includegraphics{sideview}
	\caption{Measurements and modelling. (a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling. The scale bar is \SI{100}{\micro\metre} (real size ratio). (b) Schematic side view of the cell. The constant thickness gel film (orange) is surrounded by solvent. Yellow lines show the path of transmitted light through the gel phase. A region of interest is highlighted in gray. (c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area. Crosses correspond to times in (a).}
	\label{fig:sideview}
\end{figure}

\section*{Models and scaling laws}

%Longitudinally constrained swelling of the gel film is the origin and driving force of the buckling instability. The gel area becomes in excess compared to the available area, thus the gel must fold. However, this does not explains how a wavelength is selected.

The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem. Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations~\cite{landau1986}, essentially impossible to solve in analytical form. However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}. An elastic film buckles due to excess area with respect to its boundaries. Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}. $\lambda$ is then set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$. For small deflections of  amplitude $A$, $\lambda$ scales as~\cite{Vella2009,Kolinski2009}: 
%
\begin{equation}
\lambda \sim \left( A h^3 \frac{B}{\sigma_{\perp}} \right)^{1/4}.
\label{eq:lstar}
\end{equation}

%L. Landau and E.M. Lifshitz, Theory of Elasticity (Pergamon, New York, 1986), 3rd ed.

This framework has been successfully applied to situations where $\sigma_{\perp}$ comes from the resistance of an elastic substrate ~\cite{Gough1940, Bijlaard1946}. Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection. Furthermore, (Eq.~\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.

In our situation we now have to determine $\sigma_\perp$. We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers. We confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e. with the weight acting longitudinally rather than transversely.

\begin{figure*}
	\includegraphics{Darcy_vs_Poiseuille}
	\caption{Comparing model predictions with measured wavelengths. Dots come from primary pattern, squares from secondary blisters. Lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points. Prefactors are 0.63, 0.69 and 0.67 respectively. The dashed line is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
	\label{fig:DarcyPoiseuille}
\end{figure*}

We consider two limit scenarios that leads to wrinkling patterns. In a first limit, the film of gel sits without sticking to the bottom wall of the cell, i.e. $H_1\rightarrow0$, see Supplementary Video 7. Creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$. Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_{\perp}/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_{\perp}}{h} = \frac{dA}{dt}$. Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
\begin{equation}
\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4}
\end{equation}
where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$. This expression emphasizes the kinetic nature of the wavelength selection through Darcy flow. However we note that the wavelength of a generation of the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup. Fig.~\ref{fig:growth}) once the gel is touching both walls. This arrest may be due to bending rigidity~\cite{LeGoff2014}. Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate (see Supplementary Method for the precise definition).

The opposite limiting case is when the gel film lies in the middle of the cell ($H_1=H_2$, see Supplementary Video 8) and impermeable ($\alpha\rightarrow 0$). Here destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers. By symmetry the transverse load across the gel $\sigma_{\perp}$ is also the pressure difference over the wavelength. Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_{\perp} \frac{H^3}{\eta\lambda^2}$~\cite{Poiseuille1842}. Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
\begin{equation}
\lambda_P \sim (hH)^{1/2} \Upsilon^{1/6}
\end{equation}
where we recognize the scaling for the wrinkles of an elastic film on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate. Indeed within the lubrication approximation or at low Reynolds numbers one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.

A more complete derivation of the models can be found in the Supplementary Method and yields the following prefactors:
\begin{align}
\lambda_D &= 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}\\
\lambda_P &= \pi (hH)^{1/2}\left(\frac{2}{9}\Upsilon\right)^{1/6}
\end{align}
where $1/H^3 = 1/H_1^3 + 1/H_2^3$. The crossover between the two models where $\lambda_D=\lambda_P$ is found for $H = H^* \equiv 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}$.

To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$, i.e. slow destabilisation of a highly permeable film close to a wall, the main dissipation mechanism would be the flow through the porous film; whereas for $H\gg H^*$, i.e. fast wrinkling of a poorly permeable film far from any wall, the flow in the viscous layers would dominate and set the wavelength.

\section*{Experimental tests}

For a typical gel sheet (Supplementary Table~\ref{tab:data}b) the buoyancy after syn\ae{}resis is estimated as $\Delta\rho=\SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$. This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ and therefore gravity is dismissed.

Darcy and Poiseuille models yield wavelengths that are of the same order in most of the cases under study. Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements). As summarized in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$. Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly. However we are able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones. Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$). Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Fig.~\ref{fig:generations}.

Our results are displayed in Fig.~\ref{fig:DarcyPoiseuille}. Although they provide the correct order of magnitude for $\lambda$, neither model alone is able to quantitatively reproduce the full data set. However considering only the points where $H<H^*$ (respectively $H>H^*$), the Darcy (resp. Poiseuille) model fits the data well albeit with a prefactor. Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Fig.~\ref{fig:DarcyPoiseuille}b). Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.

\section*{conclusion}

In conclusion we have taken advantage of the properties of the protein casein and carefully engineered the route to form permeable biogels that, in slit geometry, produce original cascade of wrinkling patterns when casein adhesion to the slit walls is turned off. We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flow of the solvent through the gel layer for short wavelengths. Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins. As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013}.

%role of gravity

\section*{Methods summary}

Sodium caseinate (Firmenich) is dissolved in deionised water by stirring for \SI{20}{\minute} at room temperature. The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the beginning of the experiments. We typically have about \SI{10}{\minute} to pipette the dispersion in the slit before gelation starts. Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific). Excess dye is removed by centrifugation. For confocal microscopy measurements, we add 10\% of fluorescently labelled caseins to the casein/GDL dispersion while it is fluid.

Home made optical cells are composed of a slide (RS France) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm). The optical cell is briefly heated on a hot plate so that the parafilm welds to the slide and the cover slip. Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell from \SIrange{50}{150}{\micro\metre}. The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place. To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which a polyacrylamide brush is polymerized from methacrylate groups (Euromedex).

3D data are collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation. We use either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel. Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}). Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
Data shown in Fig.~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}). The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.

\bibliographystyle{naturemag3}
\bibliography{Yaourt}

\paragraph*{Acknowledgements}
Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
The authors thank Jos\'e Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
Precious experimental ideas came from Thibaut Divoux, Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
Alan Parker at Firmenich provided the casein and GDL; 
Special thanks to Madame Biot who made M.A. Biot work accessible through the \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 
ML thanks the Region Rhône Alpes and the Programme d'Avenir Lyon - Saint Etienne (PALSE NoGELPo) for postoctoral grant. 
TG and MN acknowledge funding from the agence national de la recherche fran\c{c}aise (ANR-11-PDOC-027). 
SM and ML acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No. 258803.

\paragraph*{Author Contributions}
ML and MN conducted preliminary experiments. ML, SM and TG designed the experiments and interpreted the results. ML and TG performed quantitative experiments. ML performed data analysis and modelling. ML, SM and TG wrote the manuscript.

\paragraph*{Author Information} 
The authors declare that they have no competing financial interests. 
Correspondence and requests for materials should be addressed to TG (\href{mailto:thomas.gibaud@ens-lyon.fr}{thomas.gibaud@ens-lyon.fr}).

\clearpage
\newpage
\setcounter{figure}{0}

\section*{Supplementary material}

\subsection*{Fraction of free casein as a function of the pH}

The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL. In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for \SI{3}{\minute}. The supernatant is diluted and analysed with a UV absorption spectrometer (Ocean Optics). According to~\cite{Roefs1986}, we deduced the fraction of free casein proteins from the adsorption spectrum:
%
\begin{align}
x_{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
\text{with }A(\text{pH}) &= A_{280}-1.706 A_{320},
\end{align}
%
where $A_{280}$ is the absorption peak due to aromatic amino acids in the casein molecules plus a turbidity background and $A_{320}$ is only due to turbidity.

Above the isoelectric point, the homogeneous acidification by GDL continuously creates very small protein aggregates that are difficult to centrifuge. Such suspensions are too turbid and too rapidly evolving to reliably measure $x_{free}$.

\subsection*{Microstructure measurement}

\begin{figure}[b]
	\includegraphics{nonmonotonic}%
	\caption{Microstructure evolution. (a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row). The scale bar is \SI{10}{\micro\metre}. Arrows indicates the pore size. (b) Corresponding Fourier spectra when no adhesion on top, same times as (a). Lines are fits by Eq.~\ref{eq:fractalS}. The inset shows a schematic association between real space structures and wave numbers. (c-e) Evolution of cut-off length, susceptibility and fractal dimension in full adhesive (black line) and no adhesion on top (orange dashed line) situations. Symbols indicates the times of the images in (a).}%
	\label{fig:nonmonotonic}%
\end{figure}

To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:nonmonotonic}a). After Hann windowing we performe Fourier transform and radially average the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:nonmonotonic}b) that is fitted by the following fractal form:
\begin{equation}
I(q) = \frac{2\chi\Gamma(d)}{\left(1+\left((d+1)\xi q\right)^2\right)^{d/2}}
\label{eq:fractalS}
\end{equation}
where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension. $\Gamma$ is the gamma function. Eq.~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
%En 3D on peut citer \cite{Martin1987,Chen1986}

Figure~\ref{fig:acidification}c of the main text and Supplementary Fig.~\ref{fig:nonmonotonic}c and d show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling. The later situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope. To do so, we coat only the microscope slide with acrylamide brush. The cover slip is cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens. Neither buckling nor wrinkling is detected in such configuration.

In situation (i), the gel keeps its topology throughout the gelation and over acidification. Left of Fig.~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again. Syn\ae{}resis and swelling are thus local processes. By contrast in situation (ii), the gel as a whole compresses while forming. Since our focal plane is a fixed depth, we cannot follow the topology of the same subset of the network, however the pore size is clearly smaller than in situation (ii). The isoelectric gel, at minimum gel volume, is probably isotropic or only marginally anisotropic in $z$. This explains that subsequent swelling is not occurring only along $z$ but also in plane, causing buckling.

$\xi$ is also systematically measured in the same way at the end of each wrinkling experiment and reported in Supplementary Table~\ref{tab:data}. %Final fractal dimension was always close to 2 within error bars. Comparing $\chi$ across different setting of the microscope is not relevant.

\subsection*{Confocal images processing}

At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling). Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}. At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.

We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$). Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$. We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line. Since this criterion is local, we do not need to correct for background intensity variation. In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.

From these data, we obtain for each stack the volume and thickness of the gel phase, area of the mid surface and peak-to-peak amplitude. Velocity is obtained by time differentiation. Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.

\subsection*{Mixed Darcy--Poiseuille model}

\subsubsection*{General framework for wrinkling}

The mechanical equilibrium of a plate of thickness $h$, Young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a transverse load $\sigma_\perp$ along $z$ writes
\begin{equation}
\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_\parallel h \frac{\partial^2 w}{\partial x^2} = \sigma_\perp,
\label{eq:mecheq}
\end{equation}
where $w(x,t)$ is the deflection of the plate along $z$~\cite{Biot1957}. Since we deal with slow rates of deformation, inertia is neglected.

We decomposed the deflection in sinusoidal modes $w(x,t) =  A \sin qx$ of wavevector $q$. Eq.~(\ref{eq:mecheq}) then reads
\begin{equation}
\left(\frac{1}{12} B h^3 q^4 - \sigma_\parallel h q^2\right)w = \sigma_\perp,
\label{eq:mecheqop}
\end{equation}
where $B=E/(1-\nu^2)$ is the modulus of the plate.

Following \cite{Biot1957}, we consider the case where the compression load $\sigma_\parallel$ is increased quasistatically. The observed mode is the one that arises at the lowest possible $\sigma_\parallel$, i.e. the mode minimising $\sigma_\parallel$.

\subsubsection*{Elastic film in an infinite elastic medium}

In this framework, we recall here the case of an elastic film lying over an infinite elastic medium~\cite{Gough1940,Bijlaard1946,Biot1957,Cerda2003}. The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -q B_s w$, and thus the compression load writes
\begin{equation}
\sigma_\parallel = B \frac{h^2}{12} q^2 + \frac{B_s}{hq}.
\label{eq:sigma0}
\end{equation}

Minimizing $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
\begin{equation}
\lambda = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
\label{eq:lambdaElEl}
\end{equation}

\subsubsection*{Elastic film on a viscous layer}
Let us now turn to the case of an elastic, impermeable film on a single viscous layer. As \cite{Huang2002} we neglect gravity, although it is difficult to do so in practice at a free interface. Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation. Further considering only small deflections, we may neglect the displacement of the plate along $x$. The flux along $x$ is thus of the Poiseuille form
\begin{equation}
Q = -\frac{H^3}{12\eta}\frac{\partial p}{\partial x}.
\label{eq:PoiseuilleFlux}
\end{equation}
where $p(x)$ is the pressure in the layer of viscosity $\eta$.

Mass conservation writes
\begin{equation}
\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} = 0.
\label{eq:conservation}
\end{equation}

Assuming an exponentially growing amplitude $A(t) = A_0 \exp(t/\tau)$ and keeping only linear order terms, one gets
\begin{equation}
\frac{12\eta}{\tau} \frac{w}{H^3q^2} = -\sigma_\perp.
\label{eq:PoiseuilleLoad}
\end{equation}

Combining Eqs.~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
\begin{equation}
\sigma_\parallel = \frac{1}{12}B h^2 q^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h q^4}
\label{eq:sigma0P}
\end{equation}
and is minimized at $q_P$ such that
\begin{equation}
q_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
\label{eq:qP}
\end{equation}
with $\Upsilon = B\tau/\eta$ the viscoelastic factor. The dominant wavelength for the Poiseuille model is thus
\begin{equation}
\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
\end{equation}

We note that the mode minimising $\sigma_\parallel$ is also maximising the growth rate
\begin{equation}
\frac{1}{\tau} = \frac{H^3q^2}{12^2\eta}\left(12\sigma_\parallel - Bh^3q^4\right).
\end{equation}

When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of Eq.~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness 
\begin{equation}
\frac{1}{H^3} = \frac{1}{H_1^3}+\frac{1}{H_2^3}.
\end{equation}
This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface gravity can indeed be neglected.

\subsubsection*{Porous elastic film with no possible Poiseuille flow}
Another limiting case is that of a porous elastic film immersed in a viscous medium where longitudinal resistance to flow is infinite. It is the case in our experiments when the film is at contact with one of the walls. The only way to deform the film is to flow some liquid through the film of permeability $\alpha$. The transverse load derives from the Darcy law:
\begin{equation}
\sigma_\perp = p_1-p_2 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
\label{eq:DarcyLoad}
\end{equation}

Combining Eqs.~(\ref{eq:mecheqop}) and (\ref{eq:DarcyLoad}), the compression load writes
\begin{equation}
\sigma_\parallel = \frac{1}{12}B h^2 q^2 + \frac{\eta}{\tau}\frac{1}{\alpha q^2}
\label{eq:sigma0D}
\end{equation}
and is minimized at $q_D$ such that
\begin{equation}
q_D^4 = \frac{12}{h^2\alpha\Upsilon}.
\label{eq:qD}
\end{equation}

The dominant wavelength for Darcy model is thus
\begin{equation}
\lambda_D = 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}.
\end{equation}
Here also the dominant mode maximises the growth rate.

\subsubsection*{Porous elastic film between two viscous layers}
Finally the most realistic scenario mixes both Poiseuille and Darcy mechanisms, i.e. considers a porous elastic film between two viscous layers of arbitrary thickness. In this case, the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
\begin{equation}
H_1^3 p_1 + H_2^3 p_2 = 0.
\label{eq:pressures}
\end{equation}

The mass conservation in the lower viscous layer is the same as Eq.~(\ref{eq:conservation}) with an added leak $v$ due to the porosity
\begin{equation}
\frac{\partial Q_1}{\partial x} + \frac{\partial w}{\partial t} + v = 0.
\label{eq:conservationDarcy}
\end{equation}

$v$ can be expressed by the Darcy law
\begin{equation}
v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}.
\label{eq:Darcy}
\end{equation}

Combining Eqs.~(\ref{eq:pressures}, \ref{eq:conservationDarcy}, \ref{eq:Darcy}) we obtain the transverse load
\begin{equation}
\sigma_\perp = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 q^2}{12} + \frac{\alpha}{h}},
\label{eq:MixtLoad}
\end{equation}

Note that we recover Eq~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and Eq.~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$. We can recast this intuition in terms of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $q_P$ and $q_D$:
\begin{equation}
H^* = 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}.
\end{equation}

Analytic minimization of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome. Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
\begin{align}
\frac{12\sigma_\parallel}{q_D^2 h^2 B} &= \left(\frac{q}{q_D}\right)^2 + \frac{1}{2\frac{q_D^2}{q_P^6}q^4  + \left(\frac{q}{q_D}\right)^2}\\
&= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}
\end{align}
and we minimize this expression numerically.

\subsection*{Wavelength measurement}

\begin{figure}
	\includegraphics{generations}%
	\caption{Fourier spectra of binary altitudes of Main Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns. Inset: reconstructed binary altitudes at time \SI{2}{\hour}. The scale bar is \SI{1}{\milli\metre}.}%
	\label{fig:generations}%
\end{figure}

\begin{figure*}
	\includegraphics{patterns}
	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data} of the main text. (a-d) increasing cell thickness. (e-f) larger solid content. (g-i) increasing glycerol content. All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images. The scale is common to all panels (scale bar \SI{1}{\milli\metre}). Length of arrows correspond to the measured primary wavelength.}
	\label{fig:patterns}
\end{figure*}

To measure wrinkle wavelength, we use large field pictures of the final state. These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification (Fig.~\ref{fig:patterns} except c and e), or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x (Fig.~\ref{fig:patterns}c and e). The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film. White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall). It is thus possible to reconstruct a binary picture of the pattern (inset of Fig.~\ref{fig:generations}). Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final picture to pure first generation.

Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times. When only the first generation pattern is present a peak clearly indicates the dominant mode. When second generation has appeared, the dominant wavelength is halved and the peak much weaker. This allows us to use secondary pattern is our analysis by setting their wavelength at half of the primary wavelength. Further generations do not seem to appear simultaneously and the peak is weak and broad. This could be a signature of a transition to chaos of a non-linear oscillator~\cite{Brau2011}.

\subsection*{Characteristic time measurement}
\begin{figure}
	\includegraphics{growth}
	\caption{Growth of the instability. (a) Evolution of the gel velocity $v$ averaged over the whole surface. Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale. The black line is the best exponential fit with $\tau=\SI{90}{\second}$. (b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown). Note that the longitudinal size of the blister evolves in this regime. The thick black curve is the profile at much longer times and is at lower altitude due to swelling. Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
	\label{fig:growth}
\end{figure}

In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements. However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface. Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.

We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel. We use this approximation of $\tau$ for all our samples. This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.

\subsection*{Permeability measurements}
\begin{figure*}
	\includegraphics{permeability}
	\caption{Permeability measurements. (a) Schematic representation of the experiment. (b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2. The black line is the best exponential fit $Ae^{-t/\tau}$. (b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$. (c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
	\label{fig:permeability}
\end{figure*}

To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}. First, the gel of interest is formed at the lower extremity of a thin glass tube. The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition). Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4. The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment. Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it. A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube. Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.

Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
\begin{equation}
\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
\end{equation}
where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel. We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer. We can thus write
\begin{equation}
\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
\end{equation}

By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$. Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability. Obtaining $z_\infty$ without tube 2 would be time prohibitive.

\subsection*{Samples characteristics}
To test the theoretical predictions we have fully characterized 9 samples, see Supplementary Table \ref{tab:data}.

\begin{table*}
\begin{tabular}{cSSSScSrrScSSSScSSSSSS}
%\toprule
&\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\ 
&{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\ 
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
%ech2
a & 4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
& & & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
b & 4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
c & 4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
d & 4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
& & & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
e & 8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
& & & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
f & 8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
& & & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
g & 4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
h & 4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
i & 4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
\end{tabular}
\caption{Fully characterised samples at the basis of Fig.~\ref{fig:DarcyPoiseuille}. Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line. `cas', `GDL' and `gly' indicate the weight fraction of sodium caseinate, GDL and glycerol respectively. $\xi$ is the pore size at the end of each experiment.}
\label{tab:data}
\end{table*}

\subsection*{Supplementary Videos}

\begin{description}
\item[Video 1] Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion. Top and bottom slides of the optical cell are coated with acrylamide brushes. The duration of the Video is \SI{2}{\hour}~30. The scale bar is \SI{1}{\milli\metre}.

\item[Video 2] 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion (Supplementary Table~\ref{tab:data}c). 10\% of the proteins are fluorescently labelled. Top and bottom slides of the optical cell are coated with acrylamide brushes. The duration of the Video is \SI{2}{\hour}~30. The scale bar is \SI{1}{\milli\metre}.

\item[Video 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the Video is \SI{13}{\minute}. The scale bar is XXX $\mu$m

\item[Video 4] Titration of a dispersion with 4\%w casein by 1 molar HCl. The duration of the Video is \SI{14}{\minute}. The scale bar is XXX $\mu$m

\item[Video 5] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on. 10\% of the proteins are fluorescently labelled. The duration of the Video is \SI{132}{\minute}. The scale bar is \SI{10}{\micro\metre}.

\item[Video 6] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky. 10\% of the proteins are fluorescently labelled. The duration of the Video is \SI{147}{\minute}. The scale bar is \SI{10}{\micro\metre}.

\item[Video 7] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a). 10\% of the proteins are fluorescently labelled. The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3). Here the gel starts close to the bottom wall, $H<H^*$. Darcy mechanism dominates the formation of the primary pattern. Total duration of the video is \SI{145}{\minute}.

\item[Video 8] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}c). 10\% of the proteins are fluorescently labelled. The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3). Here the gel starts far from either top or bottom wall, $H>H^*$. Poiseuille mechanism dominates the formation of the primary pattern. Total duration of the video is \SI{181}{\minute}.

\end{description}

\end{document}