### 2016-06-08 seminaire_PIMM.tex

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage[super]{nth}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
%\usefonttheme{structuresmallcapsserif}
\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\mode<presentation>{\usetheme{ilm}}

%\usecolortheme{rose}
%%\useinnertheme[shadow]{circles}
%\usecolortheme{whale}
%%\useoutertheme{infolines}
%
%\usecolortheme[named=Accent1]{structure}
%
%\setbeamercolor{alerted text}{fg=Accent2}
%\setbeamercolor{palette primary}{fg=white}
\setbeamercolor{palette secondary}{bg=ilmcolor!10!white}
%\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}

%\setbeamerfont{page number in head/foot}{size=\large}
%\setbeamercolor{page number in head/foot}{fg=Main}
%% page/total
%%\setbeamertemplate{footline}[frame number]
%% pas de total
%\setbeamertemplate{footline}{%
%    	\hfill%
%	\usebeamercolor[fg]{page number in head/foot}%
%	\usebeamerfont{page number in head/foot}%
%	\insertframenumber\kern1em\vskip2pt%
%}
%
%\setbeamersize{text margin left=1em}
%\setbeamersize{text margin right=1em}

%font
\usepackage[T1]{fontenc}
\usepackage{times}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\makeatletter
\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
\tikzset{
  beamer externalizing/.style={%
    execute at end picture={%
      \tikzifexternalizing{%
        \ifbeamer@anotherslide
        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
        \fi
      }{}%
    }%
  },
  external/optimize=false
}
\let\orig@tikzsetnextfilename=\tikzsetnextfilename
\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
\makeatother

\tikzset{every picture/.style={beamer externalizing}}
\tikzexternalize
\tikzsetexternalprefix{fig_iLM/}
%\tikzset{external/optimize=false}
%\tikzset{external/force remake}

%link or play movies
\usepackage{multimedia}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[Lyon]{iLM, Univ. Claude Bernard Lyon 1 and CNRS, France}
\title[Yoghurt under stress]{Acid-induced protein gels: from gelation to stress-induced failure}
\author[M. Leocmach]{Mathieu Leocmach}
\date{9 June 2016}
\titlegraphic{
	\begin{tabu}{X[c]X[c]X[c]X[c]}
		\includegraphics[height=3\baselineskip]{Chris}&
		\includegraphics[height=3\baselineskip]{Thomas}&
		\includegraphics[height=3\baselineskip]{Thibaut}&
		\includegraphics[height=3\baselineskip]{Seb}\\
		Christophe Perge & Thomas\linebreak Gibaud & Thibaut\linebreak Divoux & Sebastien Manneville\\
	\end{tabu}

	\vfill
	\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
	\includegraphics[height=2\baselineskip]{CRPP}\quad
	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}
	}

\newlength{\mylength}

%\includeonly{creep_beamer}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={footnotesize}}

\pgfplotscreateplotcyclelist{earthy}{%
{red!40!black,mark=o},
{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
{red!80!black,mark=square},
{red,mark=triangle},
{red!80!yellow, mark=diamond},
red!60!yellow,
red!40!yellow,
}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}{\pgfuseimage{cnrs-logo}\hspace*{0.3cm}\pgfuseimage{ucbl-logo}}%[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{Texture as important as taste}
\vspace{1.25\baselineskip}
\tikzsetnextfilename{texture}%
\begin{tikzpicture}[inner sep=0, callout pointer segments=3]
\node(ishichan) {\includegraphics[height=8\baselineskip]{ishichan}};
\node[above=1em of ishichan] {\includegraphics[height=8\baselineskip]{calamar}};
\node[cloud callout,anchor=west, draw, callout relative pointer={(-30:2\baselineskip)}] at ($(ishichan.north)+(-0.5\textwidth-10pt,0)$) {\includegraphics[height=5\baselineskip]{aldente}};
\node[cloud callout, aspect=0.75, anchor=east, draw, callout relative pointer={(-150:2\baselineskip)}] at ($(ishichan.north)+(0.5\textwidth-10pt,0)$) {\includegraphics[height=5\baselineskip]{chewie}};
\end{tikzpicture}
\end{frame}

\begin{frame}{Mouth feel}
\begin{columns}
\column{0.8\textwidth}
\begin{itemize}
\item A very complex, multidimensional sensation
\item Behaviour of the food at large deformation?
\item How does food break into pieces?
\end{itemize}
\column{0.2\textwidth}
\includegraphics[width=\textwidth]{violon}
\end{columns}

\bigskip
\begin{columns}[T]
\column{0.475\textwidth}
\begin{block}{Ductile fracture}
\includegraphics[height=5\baselineskip]{deflated_balloon}

Irreversible deformation
\begin{itemize}
\item rubber
\item lead
\item butter
\end{itemize}
\end{block}

\column{0.5\textwidth}
\begin{block}{Brittle fracture}
\includegraphics[height=6\baselineskip]{spaghetti_breaking}\hfill
\includegraphics[height=6\baselineskip]{Tea_bowl_fixed_in_the_Kintsugi_method}
\begin{itemize}
\item concrete
\item asphalt
\item glass
\end{itemize}

\end{block}
\end{columns}
\end{frame}

%
%
%
%\begin{frame}{Fundamental issues in biogels}
%\begin{itemize}
%\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
%\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
%\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
%\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
%\end{itemize}
%\begin{tabu}{X[c]X[c]}
%\includegraphics[height=6\baselineskip]{Bonn_fracture} &
%\includegraphics[height=6\baselineskip]{Villi_sq}
%\end{tabu}
%\begin{block}{Our questions}
%\begin{itemize}
%\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
%\item What is the role of the solvent in mechanical instabilities?
%\end{itemize}
%\end{block}
%\end{frame}

%\section{Our system: Yoghurt}

%\tikzset{external/force remake}
\begin{frame}{The simplest yoghurt}
\begin{block}{Acid-set protein gel}
\begin{itemize}
\item Water (\SI{30}{\celsius})
\item Sodium caseinate (milk protein) 4\%
\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
\item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow homogeneous acidification
\end{itemize}
\end{block}

\begin{columns}
\column{0.5\textwidth}
\tikzsetnextfilename{gelation_gdl1}
\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group name=g, group size=1 by 2,
			xticklabels at=edge bottom,
			vertical sep=0,
			},
		%xmode=log,
		%xmin=1e2,xmax=3e5,
		xmin=0, xmax=20,
		scale only axis,
		width=\textwidth-4em,
		height=0.3\textwidth,
		extra tick style={grid=major},%
		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
		]
	\nextgroupplot[
		ymin=0, ymax=7, ylabel={pH},
		extra y ticks={4.6}, extra y tick labels={},%
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y190_cas4_gdl1.pH};
	\draw[help lines] (axis cs:0,4.6) -- (axis cs:3.88,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:3.88,0);

	\nextgroupplot[
		xlabel={time (\si{\hour})},
		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
		extra x ticks={17}, extra x tick labels={},
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y22.prise};
	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL1_Y22.prise};
	\begin{scope}[font=\scriptsize]
		%\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
		\node[anchor=base,rotate=80] at (axis cs:2.5,400) {gelation};
	\end{scope}
	\end{groupplot}
	%\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
\end{tikzpicture}

\column{0.5\textwidth}
\includegraphics[width=\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}

\begin{scriptsize}
\textit{Kaláb (1983),\linebreak Roefs \& van Vliet (1990),\linebreak Lucey \& Singh (1998)}
\end{scriptsize}
\end{columns}
\end{frame}
%\tikzset{external/force remake=false}

\begin{frame}{Power law linear rheology}
\[\text{stress}\rightarrow \sigma = G \gamma\leftarrow\text{strain}\]
\[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
\tikzsetnextfilename{powerlaw}
\begin{tikzpicture}
\begin{loglogaxis}[
	height=0.8\textheight,
	width=\textwidth,
	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{loss} moduli (\si{\pascal})},
	domain={6e-2:70},
	]
	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
	\node[below right] at (rel axis cs:0.1,0.9) {(visco)elastic solid};
\end{loglogaxis}
\end{tikzpicture}
\end{frame}

\section*{Creep and yielding}

\include{creep_beamer}

\begin{frame}{Can we generalise to biogels?}
\hfill\textup{\footnotesize\usebeamercolor[fg]{titlelike}(physical gels
$E_{bond}\gg k_B T$)}
\begin{tabu}{cX[c]c}
\includegraphics[height=0.3\textheight]{cell_mech} &
\includegraphics[height=0.3\textheight]{electrophoresis} &
\includegraphics[height=0.3\textheight]{cosmetics} \\
Cell mechanics & Electrophoresis & Cosmetics\\
\end{tabu}
\begin{tabu}{X[c]X[c]}
\includegraphics[height=0.3\textheight]{bacterial_culture} &
\includegraphics[height=0.3\textheight]{food} \\
Bacterial culture & Food\\
\end{tabu}

\begin{footnotesize}
\begin{block}{Sources}
\begin{tabu}{lXl}
Wikimedia Commons & www.madaboutscience.com & www.keautystore.com\\
\end{tabu}
\end{block}
\end{footnotesize}
\end{frame}

\section*{Beyond critical gels}

\begin{frame}{Acidifying casein solutions with \textup{HCl}}
\vspace{1.25\baselineskip}
\tikzsetnextfilename{dosage}%
\input{fig_presentation/dosage.pgf}
\end{frame}

\begin{frame}{Over-acidified protein gels}
	\hfill beyond isoelectric pH
	\tikzsetnextfilename{prise_cas4}
	\input{fig_soundbite_palavas/prise_cas4.pgf}
	\begin{itemize}
	\item[$\Rightarrow$] still solid
	\item From critical to super-critical gels?
	\end{itemize}
\end{frame}

\begin{frame}{Linear rheology $\rightarrow$ to soft glassy}
	\vspace{\baselineskip}
	\tikzsetnextfilename{sweep_cas4}%
\input{fig_soundbite_palavas/sweep_cas4.pgf}
	\begin{itemize}
	\item Near isoelectric, power law rheology typical of critical gels 
	\item Over-acidification: exponent $\searrow$ \hfill\textcolor{ilmorange}{\&}\hfill minimum in $G^{\prime\prime}$
	\item Transition from power-law to soft-glassy behaviour \\\textit{\footnotesize Sollich, PRE (1998); Mason, Curr. Opin. Colloid Interface Sci. (1999)}
	\end{itemize}
\end{frame}

\begin{frame}{Questions}
\begin{description}
\item[What] is the change in microstructure?
\item[How] does it affects the failure mechanism?
\end{description}
\end{frame}

\begin{frame}{Microscructure: confocal microscopy}
	\vspace{\baselineskip}
	\tikzsetnextfilename{structure_confocal}%
\input{fig_soundbite_palavas/structure_confocal.pgf}
	\vspace{-1em}
	\begin{itemize}
	\item Maximum $G^\prime$\hfill\textcolor{ilmorange}{$\leftrightarrow$}\hfill Maximum contrast between phases
	\item Pore size $\xi$ decrease during over-acidification
	\item<2> Casein aggregates swell $\sim$ over compressed $\Rightarrow$ soft glassy
	\end{itemize}
\end{frame}

\begin{frame}{3 creep regimes + softness\textsc{gdl} 4\%}
\vfill
	\tikzsetnextfilename{creep_cas4_gdl4}
	\input{fig_presentation/creep_cas4_gdl4.pgf}
\begin{itemize}
\item reversible, homogeneous creep $\rightarrow$ irreversible fracture growth
\item nucleation regime \alert{NOT universal}
\begin{itemize}
	\item possible to reform bond?
	\item friction between swollen strands?
	\item glassy dissipation?
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{Over-acidification summary}
\begin{itemize}
\item The critical network is still there but over-compressed
\item Transition to soft glassy rheology
\item Power-law creep and fracture growth still present but do not rule everything
\item[$\Rightarrow$] Rupture scenario of critical gel only partially valid for other gels.
\end{itemize}

\bigskip
\begin{center}
\structure{\huge Thank you for your attention}
\end{center}
\end{frame}

\appendix
\newcounter{finalframe}
\setcounter{finalframe}{\value{framenumber}}

\begin{frame}[plain]
\end{frame}

\begin{frame}{Creep rheology: Three regimes}
\tikzsetnextfilename{three_regimes}
\begin{tikzpicture}
\begin{loglogaxis}[
   name=g,
	width=\textwidth,
	height=0.8\textheight,
	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
	ymin=0.2,ymax=2,
	ytick={0.2, 0.4,0.8,1.6}, yticklabels={0.2, 0.4,0.8,1.6},
	ylabel={strain $\gamma$},
	extra tick style={grid=major},%
	extra y ticks={1}, extra y tick labels={1.0},
	cycle list name=earthy,
	axis on top,
	]
	\fill[Accent1!20]
		(axis cs:10,0.4) ellipse[rotate=-30, x radius=0.2\textwidth, y radius=0.12\textwidth]
		(axis cs:2e2,1.6) ellipse[x radius=0.35\textwidth, y radius=0.05\textwidth]
		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
		;
	\node[above right] at (axis cs:1e1,0.2) {primary};
	\node[below left] at (axis cs:4e5,0.7) {secondary};
	\node[left] at (axis cs:5,1.6) {tertiary};
	\addplot table{Y27_200Pa_gamma_decimated.txt} node (s200){};
	\addplot table{Y38_300Pa_gamma_decimated.txt} node (s300){};
	\addplot table{Y25_400Pa_gamma_decimated.txt}  node (s400){};
	\addplot table{Y32_550Pa_gamma_decimated.txt}  node (s550){};
	\addplot table{Y39_1000Pa_gamma_decimated.txt} node (s1000){};
\end{loglogaxis}
\begin{scope}[anchor=base, every node/.style={yshift=0.2em}]
	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
	\node[red!60!black] at (s300 |- g.outer north) {300};
	\node[red!80!black] at (s400 |- g.outer north) {400};
	\node[red] at (s550 |- g.outer north) {550};
	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
\end{scope}
\end{tikzpicture}

Failure at $\gamma\approx 1$ for a well defined time $\tau_f$
\end{frame}

\begin{frame}{Basquin law rescaling}
	\vspace{0.5\baselineskip}
	\tikzsetnextfilename{basquin_rescale}%
	\input{fig_soundbite_palavas/basquin.pgf}
\end{frame}

\begin{frame}{Monkman-Grant}
	\vspace{0.5\baselineskip}
	\hfill Predict failure from nucleation
	\tikzsetnextfilename{MonkmanGrant}
	\input{fig_soundbite_palavas/MonkmanGrant.pgf}
\end{frame}

\begin{frame}{Wavelength}
	\vspace{1.25\baselineskip}
	\tikzsetnextfilename{lambda}%
	\input{fig_soundbite_palavas/lambda.pgf}
	\textcolor{Accent1}{$\blacktriangleright$} $\lambda\propto e$ compatible with nucleation rate

\end{frame}

\setcounter{framenumber}{\value{finalframe}}
\end{document}